### 2019-05-04 paper.tex
Added sample bib and credentials

% !TeX encoding = UTF-8
% !TeX spellcheck = de_DE

%% Dies gibt Warnungen aus, sollten veraltete LaTeX-Befehle verwendet werden
\RequirePackage[l2tabu, orthodox]{nag}

\documentclass[utf8,biblatex]{lni}
\bibliography{lni-paper-example-de}

%% Schöne Tabellen mittels \toprule, \midrule, \bottomrule
\usepackage{booktabs}

%% Zu Demonstrationszwecken
\usepackage[math]{blindtext}
\usepackage{mwe}

%% BibLaTeX-Sonderkonfiguration,
%% falls man schnell eine existierende Bibliographie wiederverwenden will, aber nicht die .bib-Datei händisch anpassen möchte.
%% Bitte \iffalse und \fi entfernen, dann ist diese Konfiguration aktiviert.

\iffalse
\AtEveryBibitem{%
  \ifentrytype{article}{%
  }{%
    \clearfield{doi}%
    \clearfield{issn}%
    \clearfield{url}%
    \clearfield{urldate}%
  }%
  \ifentrytype{inproceedings}{%
  }{%
    \clearfield{doi}%
    \clearfield{issn}%
    \clearfield{url}%
    \clearfield{urldate}%
  }%
}
\fi

\begin{document}
%%% Mehrere Autoren werden durch \and voneinander getrennt.
%%% Die Fußnote enthält die Adresse sowie eine E-Mail-Adresse.
%%% Das optionale Argument (sofern angegeben) wird für die Kopfzeile verwendet.
\title[Eclipse Che for Education]{Eclipse Che for Fair and Easy Education}
%%%\subtitle{Untertitel / Subtitle} % falls benötigt
\author[Leonhard Applis]
{Leonhard Applis \footnote{Technische Hochschule Georg Simon Ohm, \email{Leonhard.Applis@Protonmail.com}}}
%\startpage{-} % Beginn der Seitenzählung für diesen Beitrag
%\editor{-}    % Namen der Herausgeber
\booktitle{Under Revision for SKILL 2019} % Name des Tagungsband
\year{2019}
%%%\lnidoi{18.18420/provided-by-editor-02} % Falls bekannt
\maketitle

\begin{abstract}
Die \LaTeX-Klasse \texttt{lni} setzt die Layout-Vorgaben für Beiträge in LNI Konferenzbänden um.
Dieses Dokument beschreibt ihre Verwendung und ist ein Beispiel für die entsprechende Darstellung.
Der Abstract ist ein kurzer Überblick über die Arbeit der zwischen 70 und 150 Wörtern lang sein und das Wichtigste enthalten sollte.
Die Formatierung erfolgt automatisch innerhalb des abstract-Bereichs.
\end{abstract}

\begin{keywords}
LNI Guidelines \and \LaTeX Vorlage
\end{keywords}

\section{Verwendung}
Die GI gibt unter \url{http://www.gi-ev.de/LNI} Vorgaben für die Formatierung von Dokumenten in der LNI Reihe.
Für \LaTeX-Dokumente werden diese durch die Dokumentenklasse \texttt{lni} realisiert.

Dieses Dokument basiert auf der offiziellen Dokumentation, simplifiziert und setzt grundlegendes LaTeX-Wissen voraus.
Es werden generische Platzhalter an die entsprechenden Stellen (wie beispielsweise die Authoren-Angaben) gesetzt und nicht weiter an anderer Stelle dokumentiert.

Dieses Template ist wie folgt gegliedert:
\Cref{sec:demos} zeigt Demonstrationen der LNI-Verlage.
\Cref{sec:lniconformance} zeigt die Einhaltung der Richtlinien durch einfachen Text.

\section{Demonstrationen}
\label{sec:demos}
Das Symbol für Potenzmengen ($\powerset$) wird korrekt angezeigt.
Es ist kein Weierstraß-p ($\wp$) mehr.

Spitze Klammen können direkt eingegeben werden: <test />

Hier eine kleine Demonstration von \href{https://www.ctan.org/pkg/microtype}{microtype}:
\blindtext

\section{Demonstration der Einhaltung der Richtlinien}
\label{sec:lniconformance}

\subsection{Literaturverzeichnis}
Der letzte Abschnitt zeigt ein beispielhaftes Literaturverzeichnis für Bücher mit einem Autor \cite{Ez10} und zwei AutorInnen \cite{AB00}, einem Beitrag in Proceedings mit drei AutorInnen \cite{ABC01}, einem Beitrag in einem LNI Band mit mehr als drei AutorInnen \cite{Az09}, zwei Bücher mit den jeweils selben vier AutorInnen im selben Erscheinungsjahr \cite{Wa14} und \cite{Wa14b}, ein Journal \cite{Gl06}, eine Website \cite{GI19} bzw.\ anderweitige Literatur ohne konkrete AutorInnenschaft \cite{XX14}.
Es wird biblatex verwendet, da es UTF8 sauber unterstützt und \href{https://github.com/gi-ev/LNI/issues/5}{im Gegensatz zu lni.bst} keine Fehler beim bibtexen auftreten.

Referenzen sollten nicht direkt als Subjekt eingebunden werden, sondern immer nur durch Authorenanganben:
Beispiel: \Citet{AB00} geben ein Beispiel, aber auch \citet{Az09}.
Hinweis: Großes C bei \texttt{Citet}, wenn es am Satzanfang steht. Dies ist analog zu \texttt{Cref}.

Formatierung und Abkürzungen werden für die Referenzen \texttt{book}, \texttt{inbook}, \texttt{proceedings}, \texttt{inproceedings}, \texttt{article}, \texttt{online} und \texttt{misc} automatisch vorgenommen.
Mögliche Felder für Referenzen können der Beispieldatei \texttt{lni-paper-example-de.bib} entnommen werden.
Andere Referenzen sowie Felder müssen allenfalls nachträglich angepasst werden.

\subsection{Abbildungen}
\Cref{fig:demo} zeigt eine Abbildung.

\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{example-image}
  \caption{Demographik}
  \label{fig:demo}
\end{figure}

\subsection{Tabellen}
\Cref{tab:demo} zeigt eine Tabelle.

\begin{table}
\centering
\begin{tabular}{lll}
\toprule
Überschriftsebenen & Beispiel & Schriftgröße und -art \\
\midrule
Titel (linksbündig) & Der Titel \ldots & 14 pt, Fett\\
Überschrift 1 & 1 Einleitung & 12 pt, Fett\\
Überschrift 2 & 2.1 Titel & 10 pt, Fett\\
\bottomrule
\end{tabular}
\caption{Die Überschriftsarten}
\label{tab:demo}
\end{table}

\subsection{Programmcode}
Die LNI-Formatvorlage verlangt die Einrückung von Listings vom linken Rand.
In der \texttt{lni}-Dokumentenklasse ist dies für die \texttt{verbatim}-Umgebung realisiert.

\begin{verbatim}
public class Hello {
    public static void main (String[] args) {
        System.out.println("Hello World!");
    }
}
\end{verbatim}

Alternativ kann auch die \texttt{lstlisting}-Umgebung verwendet werden.

\Cref{L1} zeigt uns ein Beispiel, das mit Hilfe der \texttt{lstlisting}-Umgebung realisiert ist.

\begin{lstlisting}[caption={Beschreibung}, label=L1, language=Java]
public class Hello {
    public static void main (String[] args) {
        System.out.println("Hello World!");
    }
}
\end{lstlisting}

\subsection{Formeln und Gleichungen}

Die korrekte Einrückung und Nummerierung für Formeln ist bei den Umgebungen \texttt{equation} und \texttt{eqnarray} gewährleistet.

\begin{equation}
  1=4-3
\end{equation}
und
\begin{eqnarray}
  2=7-5\\
  3=2-1
\end{eqnarray}

%% \bibliography{lni-paper-example-de.tex} ist hier nicht erlaubt: biblatex erwartet dies bei der Preambel
%% Starten Sie "biber paper", um eine Biliographie zu erzeugen.
\printbibliography

\end{document}

### 2019-05-04 paper.tex
raw transcription of introduction

@@
 \begin{abstract}
-Die \LaTeX-Klasse \texttt{lni} setzt die Layout-Vorgaben für Beiträge in LNI Konferenzbänden um.
-Dieses Dokument beschreibt ihre Verwendung und ist ein Beispiel für die entsprechende Darstellung.
-Der Abstract ist ein kurzer Überblick über die Arbeit der zwischen 70 und 150 Wörtern lang sein und das Wichtigste enthalten sollte.
-Die Formatierung erfolgt automatisch innerhalb des abstract-Bereichs.
+Eclipse Che is a revolutionary cloud-based IDE, inheriting the approach of building cloud-ready software within a 
+remote environment. 
+Instead of developing on your local machine, a remote development-server is accessed via a webbrowser 
+and the artifacts are compiled into docker-containers.
+
+This paper shortly summarizes the required environment, introduces the concept and workflow of Eclipse Che and evaluates 
+the benefits and downsides of this approach. 
+Focus is set on a new point of view, enhancing computer studies and 
+programming classes with this technology. 
 \end{abstract}
@@
 \begin{keywords}
-LNI Guidelines \and \LaTeX Vorlage
+Eclipse Che, Cloud-Ready, Cloud-Native, DevOps, Docker, Kubernetes, OpenShift
 \end{keywords}
 
-\section{Verwendung}
-Die GI gibt unter \url{http://www.gi-ev.de/LNI} Vorgaben für die Formatierung von Dokumenten in der LNI Reihe.
-Für \LaTeX-Dokumente werden diese durch die Dokumentenklasse \texttt{lni} realisiert.
+\section{Introduction}
+As children we, used to look into the clouds and count sheep. 
+As developers, we look into the clouds and check our applications.  
+Cloud technologies are already state of the art, enabling global players like Google to run applications on a world-scale.
+The market-leaders got their portfolio stocked up with  a  variety  of  different  cloud  technologies,  ranging 
+from cloud computing, cloud storage, database-as-a-service or rentable authentication. 
+Modern Software needs to be scalable, self-contained, and should run on any device, aswell as in the cloud.  
+This trend is also visible in Googles searches (TODO: Abbildung 1).
 
-Dieses Dokument basiert auf der offiziellen Dokumentation, simplifiziert und setzt grundlegendes LaTeX-Wissen voraus.
-Es werden generische Platzhalter an die entsprechenden Stellen (wie beispielsweise die Authoren-Angaben) gesetzt und nicht weiter an anderer Stelle dokumentiert.
+Figure 1(TODO LINK) shows the steadily rising interest in the most popular cloud technologies. 
+Docker is a container-engine enabling OS-Virtualization, Kubernetes is a Container-Platform manager and Openshift is a software-suite based on Kubernetes. 
+These technologies will be summarized in 2.1.  
+As cloud-technology such technologies are referred which achieve five attributes: 
+On-demand self-service, broad network-access, resource pooling, rapid elasticity and measured service [15]. 
+Cloud technology usually splits into two components:  Infrastructure  and  Services,  where  docker and Kubernetes are considered infrastructure and provide a  platform  to  host  services.  
+Most  of  the  attributes  are (pseudo-) achievable by the infrastructure alone – such as restarting containers. 
+Software running as a service can be put into two categories: 
+“Cloud-Ready” – meaning that it can be run in common cloud environments, such as Kubernetes. 
+The term cloudready  does  usually  not  include  live-scalability.
+Cloud-readieness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a dockercontainer.
+Lately the term “Cloud-native” came up [17], referring to software which is built from the very first steps to run inside the cloud.
+This kind of software is usually running inside a docker-container and both resilient and elasticby design [8].
+Resilience refers to the ability of handling failure, both from external sources and restarting/configuring themselves.  
+Elasticy is the ability to request more resources on demand, performing more computations when required, but also to free the resources if not needed. 
+Cloud-ready Software usually achieves only a certain degree of resilience.   
+Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities [16]. 
+The applications must also be able to use the gained resources in a productive and healthy way.  
+Modern  cloud-platforms  can  perform  pseudo-scalabilty for cloud-ready-services by starting multiple instances of the same service. 
+The cloud-platform performs load-balancing between these instances. 
+This naive approach can work out fine, but for example starting two services working on the same database will not yield any real scalability. 
+Another common problem arises when two services are started on the same machine and need the same port, just to name two problems with cloud-ready  software in a cloud environment.
 
-Dieses Template ist wie folgt gegliedert:
-\Cref{sec:demos} zeigt Demonstrationen der LNI-Verlage.
-\Cref{sec:lniconformance} zeigt die Einhaltung der Richtlinien durch einfachen Text.
-
+As Cloud is a central element of modern IT-development, it is strange to see that it is not yet widely teached across 
+universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
+There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud-native development is sealed behind corporate doors and conference workshops.
+Meanwhile,  the  annual  Stackoverflow  developersurvey [22] showed, that DevOps-engineers are not only 
+amongst the most wanted and most paid jobs but also tend to be the happiest participants.  
+Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career. 
+So why there are no courses? There are many reasons, but the most prominent is the complexity of the topic: 
+To build a real cloud native application, the developer must understand virtualization, system-administration, development, 
+infrastructure  and  networking. 
+Additionally, the developer needs to use specific tools, know best practices, tests and work in a team.
+All these factors make a high stake for entrance, also  represented in the developersurvey [22] which states that most DevOps engineers have a decade of experience, usually in operations topics.
+Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
+Che is unlike a normal eclipse distribution: 
+It is hosted on a server and developers get access with their browser. 
+Instead of installing dependencies, compilers etc. locally, a common workspace is set up, which is shared amongst the programmers. 
+The software build is run inside a docker-container, making it cloud-native by default. 
+With Version 7 upcoming and prominent support from RedHat, Eclipse Che is not a prototype anymore. 
+The latest distributions of RedHats OpenShift are shipped with Che in default. 
+Therefore, it is worth looking at this possible game-changer. 
+With the first part of this paper, the structure and ideas of Che are explained in more detail, the second part of this pa-
+per covers general arguments about this approach and gives a detailed overview of reachable benefits for education.    
 \section{Demonstrationen}

### 2019-05-04 paper.tex
Raw transcription of Eclipse Che

@@
 
+Figure 1
+
 Figure 1(TODO LINK) shows the steadily rising interest in the most popular cloud technologies. 
@@
 per covers general arguments about this approach and gives a detailed overview of reachable benefits for education.
+
+\section{Eclipse Che}
+\subsection{Environment and Requirements}
+Che is hosted as a server-application and is already cloudnative. 
+The only requirements are that either Docker, Kubernetes or OpenShift are available for the installation.
+The differences between these will shortly be summarized.
+
+Docker \footnote{https://www.docker.com/} is a container-plattform, where a container is a standardized unit of of software including the code, dependencies and core-functionalities. 
+A container runs on an OS-Virtualization, originating from LXC, has declared interfaces and can be parameterized. 
+One of the core features is to connect containers for bigger projects, e.g. one 
+database container and one web-server container, which are connected into a virtual network. 
+Docker is mostly famous for this docker-engine, but enriches it with monitoring and logging, as well as the ability to move containers onto different machines.
+
+Kubernetes \footnote{https://kubernetes.io/} is a container-orchestration-system based on Docker. 
+Kubernetes picked up the growing problem in managing multiple docker-containers by  simplyfing  resource-management. 
+Additional to deployment and monitoring of container-groups it provides auto-restart mechanisms and scaling based on (custom) metrics. 
+When run on Kubernetes, Che is hosted as a single docker container in the existing platform.
+The term pod, also later used in this paper, refers to a suite of connected containers. 
+An example pod would be a simple two-tier web-application, where the database and webserver are two different, but connected containers.
+
+Openshift \footnote{https://www.openshift.com/} provides a software-suite around Kubernetes with the goal to automate cloud-native development and delivery.
+Notable  additions  include  Jenkins,  Gitea,  Sonarqube and lately Che. 
+Regarding Che there are no notable changes in handling, as Che is simply run on the build in Kubernetes.
+When run on Kubernetes or Docker Che will require two additional containers for authentification with KeyCloack. 
+For Openshift the standard-openshift authentification is used.  
+Eclipse Che  can be run  on any  base-technology  on  localhost – this is rather for demonstrations aswell as for developers working on the Che-Code. 
+When run (locally) on docker, the plugin CheDir enables a portable workspace, 
+which can be used on any docker based Che.
+\subsection{Technology and Workflow}
+A normal IDE is software on your computer which makes programming easier than doing it in a plain text-editor and commandline. 
+Common tasks for an IDE are dependency management, debugging, refactoring and autocompletion.
+It is important that an IDE is not required to develop – you can write .java-files in Vim, compile on the console and run a .Jar manually. 
+With that in mind, we can see that the IDE is just a tool to make things (a lot) more comfortable – but you also have requirements such as the JDK, the .Jars you need in your classpath etc.
+as they are only utilized by your IDE.
+You can consider this stack as your workspace as it contains everything required to work, with the IDE making the workspace comfortable to use.
+Eclipse Che does not work like a normal IDE. 
+Che consists of three main components shown in Figure 2 (TODO Figure 2).
+
+FIGURE 2 CHES OVERVIEW
+
+\begin{enumerate}
+	\item Workspaces, including Runtimes and IDEs
+	\item A browser-based IDE
+	\item An administrative server
+\end{enumerate}
+A workspace is like your above mentioned common \textit{workspace}. 
+It is a single machine, as a container, containing the project files, compilers, package managers and an IDE-Interface. 
+There are two big differences to a normal development-setup:
+\begin{itemize}
+	\item The required items for the workspace are explicitly  declared, making  the  workspace itself a docker image
+	\item The IDE is not graphical – it is a REST-API performing actions like a normal IDE, such as build, writing to files and installing packages.  
+\end{itemize}
+An additional distinction is about building your software.
+Instead of building the .Jar in your remote-workspace, it is build inside a container in the remote-workspace.
+While this sounds a bit confusing, it makes sense to separate the actual runtime from the workspace. 
+This is also what happens on your machine – you compile an executable and run it separately. 
+Having these runtimes inside a docker-container comes with great benefits, outlined later in this paper.
+The second component is the web-based IDE, which is basically a webpage performing the required REST-calls and utility-tasks for the developer. 
+It is sent by the administrative server when an authorized user accesses his workspace via browser. 
+These two components can be enriched with plugins.  
+Common plugins are language-extensions which support syntax-highlighting etc. or package managers with their regarding lifecycles.
+The difference to a normal IDE plugin is that it has two components: 
+One in the workspace-API and one in the hosted web-IDE.  
+Che offers a suite of common plugins dependent on your project. 
+When starting a Maven-project, the workspace is initialized with Maven and Java.
+The last component is the Che-administration server. 
+In the administration-server the workspaces are orchestrated and monitored. 
+The authorization is also done at the admin-server and not by every workspace.
+The workspaces can be split amongst multiple machines.
+The runtimes do not necessarily need to be on the same machines as their workspaces. 
+Instead of the web-based IDE, you can connect a desktop IDE to your workspace. 
+This is done by mounting the workspaces-filesystem  in  your  desktop  IDE. 
+This  will grant you access to autocompletion and other features, but you will miss Che specific helpers the web-IDE offers. 
+My personal oppinion is, that mounting remote file systems for development sounds like a infinite source of problems. 
+I expect this feature to be rather a proof of concept.
+A single workspace can contain multiple projects, whereof every project will have its own runtime. 
+Projects can either share a versioning control or be separated. 
+You can build whole pods in a single workspace with the single containers being the subprojects. 
+You can also compose applications inside the workspaces, meaning that the resulting 
+container of a project can be the base-image of another project. 
+A workspace can be shared amongst multiple users. 
+Both will work on the same remote filesystem, altering files simultaneous. 
+Due to changes being transmitted as HTTP-requests, the current implementation when working on the same file is a \textit{last write wins}-policy. 
+For version 7 a multi-cursor file-editing will be provided, like it`s common in 
+drive-documents. 
+This is not to be mixed up with having multiple users work on the same project but on their own workspaces. 
+The code, and the workspace itself, can commonly be shared with any versioning tool like Git.
+It's mostly dependent on the team which kind of cooperation they prefer. 
+As  an  interesting  fact,  Eclipse  Che  is  developed  with Eclipse Che, proving that complex cloud development is possible. 
+This so-called \textit{dogfooding} usually leads to a product which puts requirements over any a-priori design or foreign interests.  
+Existing components will be enhanced if working with them becomes troublesome, leaving a clean and valuable core product.   
 \section{Demonstrationen}

### 2019-05-04 paper.tex
Raw transcription of General Benefits

@@
 Existing components will be enhanced if working with them becomes troublesome, leaving a clean and valuable core product.
+
+\section{General Benefits and Handicaps}
+The primary goals of Che are twofold: 
+First it reduces failure when migrating software into the cloud and enhances the experience of working in a rich remote environment.
+Second is to set up people into existing projects faster and failsafe.
+
+The university of Arizona showed [11], that about 50\% of software inspected was not able to build. 
+Common reasons are missing, changed or faulty dependencies, leading to the common term of \textit{dependency hell} aswell as unclear documentation. 
+Experienced programmers are usually able to overcome these issues, given enough time. This continues to be an issue [14], and without a different approach probably always will be. 
+This problem is solved, if the software is built inside a docker-container and follows the principles of self-containment and image-immutability [17].
+The self-containment principle dictates that  at build-time every required dependency is given, and a fully functional image is created.
+A dockerized maven-application would include java, maven and every required maven package to run the application in the current state. 
+The image-immutability means to forbid alternations to the environment. 
+No installations at runtime, no change in ports, no creation of database-schemas.
+These two principles, while being optional, lead to always reproducible and functional artifacts.
+The main challenge is to migrate existing software and fulfill these properties. 
+Typically, after making the first iterations of your software on your local machine, dockering it will yield to 100 problems at once, not only being frustrating but taking tons of time. 
+With Che, this kind of Big-Bang-Migration is not possible, as from the very first step you are working inside of a docker-container.
+The properties above can be hold with every change to the image and checked in a normal review-process. 
+Problems will arise one at a time and be solved one at a time, whether they are about code or operations. 
+Che therefore embraces the core-ideas of DevOps [19][20] and agile software development [18].
+That being said, of course it's possible to big-bang-migrate into Che, causing the normal problem mentioned above. 
+However, once the errors are resolved, every further development will work more agile by design.
+The artifacts (whether containers or pods) produced with Che are self-contained and immutable, therefore every result is reproducible. 
+Any further development is incremental on the existing images and keeps old artifacts stable. 
+As the code is fully build inside containers and are already run remotely, the famous “It works on my Machine!” is eliminated. 
+This does not only enhance deployments, but also has a big impact on quality assurance and the motivation of developers. 
+You can keep multiple projects in a single workspace and launch multiple pods at once, running the full environment. 
+The behavior of this multipod is unified and testable, making a stable release of full environments possible. 
+This is one of the primary goals of the DevOps approach [20]. 
+The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{he Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of microservices [21], while being agile and reproducible.
+Onboarding new teammembers is made easy with Che. 
+A newly joining member does not need to install IDEs, languages, tools and packages, he only needs access to his Che-Account and a singular workspace with every dependency build for him.
+Depending on the complexitiy of the project this required atmost an hour. 
+While this may seem long, the setup time is dependent on the complexity of the workspace. 
+Setting up a complex workplace will also take more time manually. 
+The work for this is done automatic and remote, so the user is free to dive into the documentation or have some coffee. 
+His workspace will be able to launch the latest version immidieatly.
+It is also notable, that the (technical) onboarding of developers and operators is streamlined into one single process. 
+The combination of a quick setup and the usability of every enhancement makes Che especially interesting to recruit troubleshooters into your projects, which solve some of the very specialiced tasks.
+With  the  artifacts  being  reproducible,  the  expert  can quickly inspect the subject and analyse problems.
+With a finished workspace he can quickly start working. 
+With his changes being also in docker, they will be visible and functional for the whole team.
+Therefore,  the  hardest  common  problems  regarding \textit{quick help} are addressed and solved. 
+As an additional minor benefit, it is easy to hibernate resources  using  Che  with  Kubernetes  or  Openshift.
+The workspaces collect metrics like any other container and are therefore scalable on demand. 
+The ability of downscaling the development resources has a big impact on the overall 
+workload. 
+As a further example, the workspaces could be shut down out of normal business-time. This attribute is crucial for a rented infrastructure, which is paid commonly on both workload and storage.  
+The very first handicap coming with Che is the requirement for a good internet connection. 
+While your normal buro is already equipped with internet, one may want to work on train or there is simply a problem with the WiFi in your hotel.
+With the possibility to setup che on localhost, you can learn the basics and routines offline. 
+As the main benefits come in terms of teamwork, this is not a real solution.
+The primary problem with Che is the requirement for an already dense knowledge of cloud related technology.
+Setting up the Workspace, docker-images, build and CI scripts is a complex topic and can not be done by someone getting started. 
+This requirement extends when working with a multipod-environment. 
+While most of the tasks are by themselves rather simple, the bare amount of such tasks (especially when migrating an existing project to Che) requires a broad knowledge.
+This can be done by a single expert user, enabling the other developers  to  use  the  build-routines and  environments provided. 
+However, many problems which arise are related to the cloud-environement and will be hard to solve by a novice user, making him dependent on the expert. 
+It's therefore necessary to get everyone on a certain base-level to solve their own problems, otherwise the expert will be overloaded and the project stuck.  
+Additionally, the missing default “build and run” is a paradigm-shift for many developers, which will take a while to get used to. 
+Clark et al. [14] describe this rather as a change in the mindset, than in actual technology:
+
+TODO CITE CITE CITE
+The primary shift that’s required is not one of new tooling, as most developers already have the basic tooling they need. 
+Rather, the needed shift is one of philosophy. 
+TODO CITE CITE CITE END
+
+Intuitively this is true: Every developer needs to setup his own workspace – the skills to do this are already there. 
+The only difference is to setup the workspace in a common, documentated and descriptive way.
+
+Another problem with Che is the uncommon debugging. 
+As the code is running in a remote-runtime, the common breakpoint  and  valuecheck
+is  missing.  
+Instead,  one  either must be familiar with the languages commandline-debugging or write good logs and metrics. 
+While this will slow down development, it also offers some benefits: 
+The debugging and error-analysis inside the container is the same as it would be in the productive environment. 
+Having meaningful logs with different loglevels and collecting valuable metrics is 
+in general a good attribute for software in any circumstance. 
+It just means additional effort and know-how.  
+The last problem is the vulnerability to malevolent users. 
+Someone who already is in possession of a valid account can drain the shared resources of a full cluster, therefore blocking  any  other  user  from  work.
+This  can  also  be achieved by human error. 
+Dependent on the setup, the attacker could also auto-deploy hidden spyware into a pro-
+duction environment. 
+While this is also a problem for normal developer teams, the fast pace of deployments with Che requires additional care. 
+A rich CI/CD environment, such as a well-maintained Jenkins may lead to less-detailed code-reviews, as both unit- and end-to-end-tests succeed.  
+It is therefore important to not only restrict access to project resources, but also the double-check and guard the releases.  
+Some security risks can also emerge from docker-images as well [9] but are out of scope for this paper.
+
 \section{Demonstrationen}

### 2019-05-04 paper.tex
Raw transcription of Educational Benefits

@@
 
+\section{Benefits for Education}
+In addition to the previous mentioned benefits, there are more points about Che that are especially important for educational purposes, such as courses in the university, professional school or research departments.
+For simplicity I will only elaborate universities, to which I can best relate.
+Eclipse Che provides the possibility to learn primary cloud-technologies hands on and in a simple and supportive manner. 
+Unfortunetally, these topics are currently not widely teached at university.
+Che makes it possible to teach students fundamentals of modern software development.
+Depending on the courses subject, one can either provide scalable software or a running environment. 
+With fitting material provided, it's easy to learn cloud-native development, DevOps or Cloud-Operations. 
+Even complicated topics, like microservices and IoT-applications can easily be adaptet inside Che [21] and performed as for example as group-projects with a common goal. 
+The student will need to understand cloud-components aswell as team-work to succeed with Che. 
+The produced artifacts can simply be shipped as containers, giving comfortable access to the professor. 
+With both the product and the code inside the workspace easily accessible, the professor can grant faster and better feedback for the students, which can have massive impact on learning [1][5].
+Additionally, the professor can provide automated end-to-end tests with tools like Jenkins or TravisCI. 
+Given such an environment, the student will immideatly get objective feedback, whether their code fulfills the requirements. 
+This immediate feedback is a big motivator and has a positive impact on the learning experience [2]. 
+With a clear reason on the usefulness of the task, the visible progress and the direct impact of personal effort should lead to intrinsic motivation, which studies show to have a strong effect on education [23].
+As a requirement, it's necessary to provide a meaningful and fun task. 
+Fortunately, such tasks are not that rare in development.  
+Other studies have shown that only automated and unified testing is not beneficial to education in general [3]. 
+But the tools provided by Che and CI/CD slim down the overhead until both personal feedback and automated can be supplied. 
+If the whole classroom utilizes this system, the educator can also monitor who is having trouble in an automated manner. 
+With a single investment for the tests, feedback for every student is granted at every time. 
+With the single investment of providing a running Che instance (and a short lecture how to use it), feedback on the artifacts and code can be immidieatly granted to every student, without having the overhead of setting up workspaces.
+These investments can be done before the course starts at any time and can be reused.
+It's not dependent on the individual students. 
+In general, this approach enables studies to be less about the subject, but instead to be about learning the subject. 
+I consider this the primary goal of university courses. 
+Another benefit for education is about fairness in IT-Education.  
+With access to a university Che workspace, every person can develop the same software in the same environment. 
+A wellknown problem is the necessity of an IOS-build-device for IOS-appdevelopment.
+Students which are limited in their access to the university-laboratory or cannot afford a mac are therefore handicapped.
+Another  rising  problem  are  GPU-accelerated  machine learning algorithms. 
+Depending on the task, training neural networks can be 20 times faster if a CUDA-compatible graphics card is used. 
+The range for these graphiccards start at 300\$ and are therefore not commodity hardware,especially for students.  
+While these technologies are great and important for many applications, the required technologies are exclusive due to  their  costs.
+This  hindrance  for  education  is  usually adressed  with  IOS-laboratories  and/or  cloud-resources paid for by university. 
+The topic of socio-economic factors is beyond the scope of this work, but the impact has been analyzed by the PISA Studies [7], which state that socio-economicly handicapped students fail three times more likely to achieve a basic level of performance.
+According to PISA, this rate has not significally changed since the studies started in 2006. 
+Germany scored on the impact of socioeconomic-factors on education better than it did in 2006 but is worse than global average [6].
+The Mac-Laboratories and rented cloudspace are technically a good attempt at inclusion. 
+But students which are required to get to university to do their assignments are 
+less  likely  to  do  so.
+Making  assignments  mandatory, providing extrinsic motivation [23] has usually a negative impact on the subjective experience.
+A different problem comes with machine-learning, which is usually a process of trial and error to some extend.
+Long waiting times disconnect the action from the result being not beneficial for the learning process [1][23]. 
+One may argue that having this kind of anticipation is extra rewarding and therefore highly motivational.
+From my personal experience, working in this modus becomes exhausting after a period of weeks, especially when errors occur. 
+I expect this experience to be highly personal. 
+I think providing the fastest infrastructure possible, enabling the fastest results, is the better way. 
+Making progress slower is much easier than making it faster.  
+In addition, the time spent waiting on your machine learning algorithm or moving to the mac-laboratory on weekends is simply dead time. 
+It should therefore be minimized.   
+With studies showing basics of good learning experiences [2][3] and the impact of motivation [1][23], it is mandatory for institutions to provide the best possible environment. 
+While many socio-economic factors cannot be grasped at the university or by the educators, those originating in the students' economic background can be addressed.
+Most universities already provide free services, such as software-suites, e-libraries or cloud storage.
+Simply providing a technical correct solution (like laboratories) is not enough to provide a good environment.
+One student with a better socio-economic background who can afford a 300\$ GPU will have a better baseline for his education. 
+This is simply put unfair.
+With Che, universities can provide a free, good and accessible development environment for every student.
+
 \section{Demonstrationen}

### 2019-05-04 paper.tex
Raw transcription of Conclusion and Future Work

@@
 With Che, universities can provide a free, good and accessible development environment for every student.
-
+\section{Conclusion}
+Cloud technology will be dominant in the foreseeable future. 
+Not \textit{everything} will be best fit for cloud, but most of the everyday software benefits from the concepts. 
+With more and more tools making it easier, it's just a matter of time until cloud-tec becomes standard instead of bonus.
+With Eclipse Che many problems about teaching students these Cloud Skills and reducing the initial scope for learning, are addressed. 
+While the impacts on motivation and the actual learning results are only hypothetical, most of the technical aspects are solved by Che.  
+As Eclipse Che is cutting edge it whatsoever lacks many quality of life features, has some bugs, and the design is undefined to some extend. 
+It somewhat is a chance: 
+Instances like a university can have massive impact on the design of Che and drive the project to some extend for their own needs. 
+With faculties working on the open source project, either providing code and features, or just testing Che in educational environments, both quality issues and missing features can be solved.    
+Now cloud development is only for corporations. 
+There is hardly any chance to successfully learn it at university. 
+With Che there is another chance to “get back into the game” – and provide corporation-free education.  
+While this sounds harsh towards companies, I think unbiase education for students is not a nice thing to have, but a core concept of education in general. 
+Only if that is true, the students can make unbiased decisions and make unbiased research. 
+Being dependent on Google for Universities is in a similar way disturbing, as if a biology faculty would be sponsored by Bayer. 
+One can only imagine a grim future if the curriculum is influenced by profit. 
+\section*{Future Work}
+As this paper only summarizes benefits and outlines some ideas, the next step would be to put it to the test.
+In addition, a study could be elaborated on both the acceptance and effect of a CI/CD based development course with Che. 
+Unfortunately, the study would lack any comparison as there are few courses on this topic. 
+These studies could be done with universities and corporations, which 
+may yield very different results.   
 \section{Demonstrationen}

### 2019-05-04 paper.tex
Removed Dummy Latex

@@
 may yield very different results.
-\section{Demonstrationen}
-\label{sec:demos}
-Das Symbol für Potenzmengen ($\powerset$) wird korrekt angezeigt.
-Es ist kein Weierstraß-p ($\wp$) mehr.
-
-Spitze Klammen können direkt eingegeben werden: <test />
-
-Hier eine kleine Demonstration von \href{https://www.ctan.org/pkg/microtype}{microtype}:
-\blindtext
-
-\section{Demonstration der Einhaltung der Richtlinien}
-\label{sec:lniconformance}
-
-\subsection{Literaturverzeichnis}
-Der letzte Abschnitt zeigt ein beispielhaftes Literaturverzeichnis für Bücher mit einem Autor \cite{Ez10} und zwei AutorInnen \cite{AB00}, einem Beitrag in Proceedings mit drei AutorInnen \cite{ABC01}, einem Beitrag in einem LNI Band mit mehr als drei AutorInnen \cite{Az09}, zwei Bücher mit den jeweils selben vier AutorInnen im selben Erscheinungsjahr \cite{Wa14} und \cite{Wa14b}, ein Journal \cite{Gl06}, eine Website \cite{GI19} bzw.\ anderweitige Literatur ohne konkrete AutorInnenschaft \cite{XX14}.
-Es wird biblatex verwendet, da es UTF8 sauber unterstützt und \href{https://github.com/gi-ev/LNI/issues/5}{im Gegensatz zu lni.bst} keine Fehler beim bibtexen auftreten.
-
-Referenzen sollten nicht direkt als Subjekt eingebunden werden, sondern immer nur durch Authorenanganben:
-Beispiel: \Citet{AB00} geben ein Beispiel, aber auch \citet{Az09}.
-Hinweis: Großes C bei \texttt{Citet}, wenn es am Satzanfang steht. Dies ist analog zu \texttt{Cref}.
-
-Formatierung und Abkürzungen werden für die Referenzen \texttt{book}, \texttt{inbook}, \texttt{proceedings}, \texttt{inproceedings}, \texttt{article}, \texttt{online} und \texttt{misc} automatisch vorgenommen.
-Mögliche Felder für Referenzen können der Beispieldatei \texttt{lni-paper-example-de.bib} entnommen werden.
-Andere Referenzen sowie Felder müssen allenfalls nachträglich angepasst werden.
-
-\subsection{Abbildungen}
-\Cref{fig:demo} zeigt eine Abbildung.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\textwidth]{example-image}
-  \caption{Demographik}
-  \label{fig:demo}
-\end{figure}
-
-\subsection{Tabellen}
-\Cref{tab:demo} zeigt eine Tabelle.
-
-\begin{table}
-\centering
-\begin{tabular}{lll}
-\toprule
-Überschriftsebenen & Beispiel & Schriftgröße und -art \\
-\midrule
-Titel (linksbündig) & Der Titel \ldots & 14 pt, Fett\\
-Überschrift 1 & 1 Einleitung & 12 pt, Fett\\
-Überschrift 2 & 2.1 Titel & 10 pt, Fett\\
-\bottomrule
-\end{tabular}
-\caption{Die Überschriftsarten}
-\label{tab:demo}
-\end{table}
-
-\subsection{Programmcode}
-Die LNI-Formatvorlage verlangt die Einrückung von Listings vom linken Rand.
-In der \texttt{lni}-Dokumentenklasse ist dies für die \texttt{verbatim}-Umgebung realisiert.
-
-\begin{verbatim}
-public class Hello {
-    public static void main (String[] args) {
-        System.out.println("Hello World!");
-    }
-}
-\end{verbatim}
-
-Alternativ kann auch die \texttt{lstlisting}-Umgebung verwendet werden.
-
-\Cref{L1} zeigt uns ein Beispiel, das mit Hilfe der \texttt{lstlisting}-Umgebung realisiert ist.
-
-\begin{lstlisting}[caption={Beschreibung}, label=L1, language=Java]
-public class Hello {
-    public static void main (String[] args) {
-        System.out.println("Hello World!");
-    }
-}
-\end{lstlisting}
-
-\subsection{Formeln und Gleichungen}
-
-Die korrekte Einrückung und Nummerierung für Formeln ist bei den Umgebungen \texttt{equation} und \texttt{eqnarray} gewährleistet.
-
-\begin{equation}
-  1=4-3
-\end{equation}
-und
-\begin{eqnarray}
-  2=7-5\\
-  3=2-1
-\end{eqnarray}
 

### 2019-05-04 paper.tex
Transcribed sources

@@
 With Che there is another chance to “get back into the game” – and provide corporation-free education.  
-While this sounds harsh towards companies, I think unbiase education for students is not a nice thing to have, but a core concept of education in general. 
+While this sounds harsh towards companies, I think unbiased education for students is not a nice thing to have, but a core concept of education in general. 
 Only if that is true, the students can make unbiased decisions and make unbiased research. 

### 2019-05-04 paper.tex
used new bib in Tex
Added Images, labels

@@
 \documentclass[utf8,biblatex]{lni}
-\bibliography{lni-paper-example-de}
+\bibliography{che}
 
@@
 \section{Introduction}
+\label{sec:Introduction}
 As children we, used to look into the clouds and count sheep. 
@@
 Modern Software needs to be scalable, self-contained, and should run on any device, aswell as in the cloud.  
-This trend is also visible in Googles searches (TODO: Abbildung 1).
+This trend is also visible in Googles searches \ref{fig:digdocker}.
 
-Figure 1
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\linewidth]{imgs/DigDocker}
+	\caption{Rising interest in Cloud Tec, Data provided by Google Trends}
+	\label{fig:digdocker}	
+\end{figure}
 
-Figure 1(TODO LINK) shows the steadily rising interest in the most popular cloud technologies. 
+Figure \ref{fig:digdocker} shows the steadily rising interest in the most popular cloud technologies. 
 Docker is a container-engine enabling OS-Virtualization, Kubernetes is a Container-Platform manager and Openshift is a software-suite based on Kubernetes. 
-These technologies will be summarized in 2.1.  
+These technologies will be summarized in \ref{subsec:Req}.  
 As cloud-technology such technologies are referred which achieve five attributes: 
@@
 Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career. 
-So why there are no courses? There are many reasons, but the most prominent is the complexity of the topic: 
+So why are there no courses? There are many reasons, but the most prominent is the complexity of the topic: 
 To build a real cloud native application, the developer must understand virtualization, system-administration, development, 
@@
 \section{Eclipse Che}
+\label{sec:Che}
 \subsection{Environment and Requirements}
+\label{subsec:Req}
 Che is hosted as a server-application and is already cloudnative. 
@@
 
-Docker \footnote{https://www.docker.com/} is a container-plattform, where a container is a standardized unit of of software including the code, dependencies and core-functionalities. 
+Docker\footnote{https://www.docker.com/} is a container-plattform, where a container is a standardized unit of of software including the code, dependencies and core-functionalities. 
 A container runs on an OS-Virtualization, originating from LXC, has declared interfaces and can be parameterized. 
@@
 
-Kubernetes \footnote{https://kubernetes.io/} is a container-orchestration-system based on Docker. 
+Kubernetes\footnote{https://kubernetes.io/} is a container-orchestration-system based on Docker. 
 Kubernetes picked up the growing problem in managing multiple docker-containers by  simplyfing  resource-management. 
@@
 
-Openshift \footnote{https://www.openshift.com/} provides a software-suite around Kubernetes with the goal to automate cloud-native development and delivery.
+Openshift\footnote{https://www.openshift.com/} provides a software-suite around Kubernetes with the goal to automate cloud-native development and delivery.
 Notable  additions  include  Jenkins,  Gitea,  Sonarqube and lately Che. 
@@
 \subsection{Technology and Workflow}
+\label{subsec:Tec}
 A normal IDE is software on your computer which makes programming easier than doing it in a plain text-editor and commandline. 
@@
 Eclipse Che does not work like a normal IDE. 
-Che consists of three main components shown in Figure 2 (TODO Figure 2).
+Che consists of three main components shown in Figure \ref{fig:cheworkflow}\footnote{source:\url{https://en.wikipedia.org/wiki/Eclipse_Che#/media/File:Eclipse_Che_-_Workflow.PNG}}.
 
-FIGURE 2 CHES OVERVIEW
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\linewidth]{imgs/Che_Workflow}
+	\caption[Che Workflow]{Che Workflow}
+	\label{fig:cheworkflow}
+\end{figure}
 
@@
 \section{General Benefits and Handicaps}
+\label{sec:Eval}
 The primary goals of Che are twofold: 
@@
 \section{Benefits for Education}
+\label{sec:EvalEdu}
 In addition to the previous mentioned benefits, there are more points about Che that are especially important for educational purposes, such as courses in the university, professional school or research departments.
@@
 \section{Conclusion}
+\label{sec:Conclusion}
 Cloud technology will be dominant in the foreseeable future. 
@@
 \section*{Future Work}
+\label{sec:FutureWork}
 As this paper only summarizes benefits and outlines some ideas, the next step would be to put it to the test.

### 2019-05-04 paper.tex
Re-Implemented Citing

@@
 As cloud-technology such technologies are referred which achieve five attributes: 
-On-demand self-service, broad network-access, resource pooling, rapid elasticity and measured service [15]. 
+On-demand self-service, broad network-access, resource pooling, rapid elasticity and measured service \cite{ME11}. 
 Cloud technology usually splits into two components:  Infrastructure  and  Services,  where  docker and Kubernetes are considered infrastructure and provide a  platform  to  host  services.  
@@
 Cloud-readieness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a dockercontainer.
-Lately the term “Cloud-native” came up [17], referring to software which is built from the very first steps to run inside the cloud.
-This kind of software is usually running inside a docker-container and both resilient and elasticby design [8].
+Lately the term “Cloud-native” came up \cite{BI17}, referring to software which is built from the very first steps to run inside the cloud.
+This kind of software is usually running inside a docker-container and both resilient and elasticby design \cite{TO17}.
 Resilience refers to the ability of handling failure, both from external sources and restarting/configuring themselves.  
@@
 Cloud-ready Software usually achieves only a certain degree of resilience.   
-Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities [16]. 
+Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities \cite{MIN18}. 
 The applications must also be able to use the gained resources in a productive and healthy way.  
@@
 There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud-native development is sealed behind corporate doors and conference workshops.
-Meanwhile,  the  annual  Stackoverflow  developersurvey [22] showed, that DevOps-engineers are not only 
+Meanwhile,  the  annual  Stackoverflow  developersurvey \cite{SO19} showed, that DevOps-engineers are not only 
 amongst the most wanted and most paid jobs but also tend to be the happiest participants.  
@@
 Additionally, the developer needs to use specific tools, know best practices, tests and work in a team.
-All these factors make a high stake for entrance, also  represented in the developersurvey [22] which states that most DevOps engineers have a decade of experience, usually in operations topics.
+All these factors make a high stake for entrance, also  represented in the developersurvey \cite{SO19} which states that most DevOps engineers have a decade of experience, usually in operations topics.
 Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
@@
 Eclipse Che does not work like a normal IDE. 
-Che consists of three main components shown in Figure \ref{fig:cheworkflow}\footnote{source:\url{https://en.wikipedia.org/wiki/Eclipse_Che#/media/File:Eclipse_Che_-_Workflow.PNG}}.
+Che consists of three main components shown in Figure \ref{fig:cheworkflow}\footnote{source: \url{https://en.wikipedia.org/wiki/Eclipse_Che#/media/File:Eclipse_Che_-_Workflow.PNG}}.
 
@@
 
-The university of Arizona showed [11], that about 50\% of software inspected was not able to build. 
+The university of Arizona showed \cite{CO16}, that about 50\% of software inspected was not able to build. 
 Common reasons are missing, changed or faulty dependencies, leading to the common term of \textit{dependency hell} aswell as unclear documentation. 
-Experienced programmers are usually able to overcome these issues, given enough time. This continues to be an issue [14], and without a different approach probably always will be. 
-This problem is solved, if the software is built inside a docker-container and follows the principles of self-containment and image-immutability [17].
+Experienced programmers are usually able to overcome these issues, given enough time. This continues to be an issue \cite{CL14}, and without a different approach probably always will be. 
+This problem is solved, if the software is built inside a docker-container and follows the principles of self-containment and image-immutability \cite{BI17}.
 The self-containment principle dictates that  at build-time every required dependency is given, and a fully functional image is created.
@@
 Problems will arise one at a time and be solved one at a time, whether they are about code or operations. 
-Che therefore embraces the core-ideas of DevOps [19][20] and agile software development [18].
+Che therefore embraces the core-ideas of DevOps \cite{HUE12}\cite{JA16} and agile software development \cite{EBE16}.
 That being said, of course it's possible to big-bang-migrate into Che, causing the normal problem mentioned above. 
@@
 The behavior of this multipod is unified and testable, making a stable release of full environments possible. 
-This is one of the primary goals of the DevOps approach [20]. 
-The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{he Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of microservices [21], while being agile and reproducible.
+This is one of the primary goals of the DevOps approach \cite{JA16}. 
+The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{he Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of microservices \cite{IV18}, while being agile and reproducible.
 Onboarding new teammembers is made easy with Che. 
@@
 Additionally, the missing default “build and run” is a paradigm-shift for many developers, which will take a while to get used to. 
-Clark et al. [14] describe this rather as a change in the mindset, than in actual technology:
+Clark et al. \cite{CL14} describe this rather as a change in the mindset, than in actual technology:
 
@@
 It is therefore important to not only restrict access to project resources, but also the double-check and guard the releases.  
-Some security risks can also emerge from docker-images as well [9] but are out of scope for this paper.
+Some security risks can also emerge from docker-images as well \cite{PE13} but are out of scope for this paper.
 
@@
 With fitting material provided, it's easy to learn cloud-native development, DevOps or Cloud-Operations. 
-Even complicated topics, like microservices and IoT-applications can easily be adaptet inside Che [21] and performed as for example as group-projects with a common goal. 
+Even complicated topics, like microservices and IoT-applications can easily be adaptet inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
 The student will need to understand cloud-components aswell as team-work to succeed with Che. 
 The produced artifacts can simply be shipped as containers, giving comfortable access to the professor. 
-With both the product and the code inside the workspace easily accessible, the professor can grant faster and better feedback for the students, which can have massive impact on learning [1][5].
+With both the product and the code inside the workspace easily accessible, the professor can grant faster and better feedback for the students, which can have massive impact on learning \cite{HU95}\cite{HU18}.
 Additionally, the professor can provide automated end-to-end tests with tools like Jenkins or TravisCI. 
 Given such an environment, the student will immideatly get objective feedback, whether their code fulfills the requirements. 
-This immediate feedback is a big motivator and has a positive impact on the learning experience [2]. 
-With a clear reason on the usefulness of the task, the visible progress and the direct impact of personal effort should lead to intrinsic motivation, which studies show to have a strong effect on education [23].
+This immediate feedback is a big motivator and has a positive impact on the learning experience \cite{FI05}. 
+With a clear reason on the usefulness of the task, the visible progress and the direct impact of personal effort should lead to intrinsic motivation, which studies show to have a strong effect on education \cite{DECI93}.
 As a requirement, it's necessary to provide a meaningful and fun task. 
 Fortunately, such tasks are not that rare in development.  
-Other studies have shown that only automated and unified testing is not beneficial to education in general [3]. 
+Other studies have shown that only automated and unified testing is not beneficial to education in general \cite{TC88}. 
 But the tools provided by Che and CI/CD slim down the overhead until both personal feedback and automated can be supplied. 
@@
 This  hindrance  for  education  is  usually adressed  with  IOS-laboratories  and/or  cloud-resources paid for by university. 
-The topic of socio-economic factors is beyond the scope of this work, but the impact has been analyzed by the PISA Studies [7], which state that socio-economicly handicapped students fail three times more likely to achieve a basic level of performance.
+The topic of socio-economic factors is beyond the scope of this work, but the impact has been analyzed by the PISA Studies \cite{PISA15}, which state that socio-economicly handicapped students fail three times more likely to achieve a basic level of performance.
 According to PISA, this rate has not significally changed since the studies started in 2006. 
-Germany scored on the impact of socioeconomic-factors on education better than it did in 2006 but is worse than global average [6].
+Germany scored on the impact of socioeconomic-factors on education better than it did in 2006 but is worse than global average \cite{OECDCN15}.
 The Mac-Laboratories and rented cloudspace are technically a good attempt at inclusion. 
@@
 less  likely  to  do  so.
-Making  assignments  mandatory, providing extrinsic motivation [23] has usually a negative impact on the subjective experience.
+Making  assignments  mandatory, providing extrinsic motivation \cite{DECI93} has usually a negative impact on the subjective experience.
 A different problem comes with machine-learning, which is usually a process of trial and error to some extend.
-Long waiting times disconnect the action from the result being not beneficial for the learning process [1][23]. 
+Long waiting times disconnect the action from the result being not beneficial for the learning process \cite{HU95}\cite{DECI93}. 
 One may argue that having this kind of anticipation is extra rewarding and therefore highly motivational.
@@
 It should therefore be minimized.   
-With studies showing basics of good learning experiences [2][3] and the impact of motivation [1][23], it is mandatory for institutions to provide the best possible environment. 
+With studies showing basics of good learning experiences \cite{FI05}\cite{TC88} and the impact of motivation \cite{HU95}\cite{DECI93}, it is mandatory for institutions to provide the best possible environment.
 While many socio-economic factors cannot be grasped at the university or by the educators, those originating in the students' economic background can be addressed.

### 2019-05-04 che.tex
Added Bib, compiling is bad :(
Subsections and paragraphs

--- 2019-05-04 paper.tex
@@
 % !TeX encoding = UTF-8
-% !TeX spellcheck = de_DE
+% !TeX spellcheck = en_EN
 
-%% Dies gibt Warnungen aus, sollten veraltete LaTeX-Befehle verwendet werden
-\RequirePackage[l2tabu, orthodox]{nag}
-
-\documentclass[utf8,biblatex]{lni}
-\bibliography{che}
-
-%% Schöne Tabellen mittels \toprule, \midrule, \bottomrule
-\usepackage{booktabs}
-
-%% Zu Demonstrationszwecken
-\usepackage[math]{blindtext}
-\usepackage{mwe}
-
-%% BibLaTeX-Sonderkonfiguration,
-%% falls man schnell eine existierende Bibliographie wiederverwenden will, aber nicht die .bib-Datei händisch anpassen möchte.
-%% Bitte \iffalse und \fi entfernen, dann ist diese Konfiguration aktiviert.
-
-\iffalse
-\AtEveryBibitem{%
-  \ifentrytype{article}{%
-  }{%
-    \clearfield{doi}%
-    \clearfield{issn}%
-    \clearfield{url}%
-    \clearfield{urldate}%
-  }%
-  \ifentrytype{inproceedings}{%
-  }{%
-    \clearfield{doi}%
-    \clearfield{issn}%
-    \clearfield{url}%
-    \clearfield{urldate}%
-  }%
-}
-\fi
+\documentclass[utf8]{lni}
 
@@
 \begin{keywords}
-Eclipse Che, Cloud-Ready, Cloud-Native, DevOps, Docker, Kubernetes, OpenShift
+Eclipse Che \and Cloud-Ready \and Cloud-Native \and DevOps \and Docker \and Kubernetes \and OpenShift
 \end{keywords}
@@
 These technologies will be summarized in \ref{subsec:Req}.
+
 As cloud-technology such technologies are referred which achieve five attributes: 
@@
 Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career.
+
 So why are there no courses? There are many reasons, but the most prominent is the complexity of the topic: 
@@
 All these factors make a high stake for entrance, also  represented in the developersurvey \cite{SO19} which states that most DevOps engineers have a decade of experience, usually in operations topics.
+
 Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
@@
 Therefore, it is worth looking at this possible game-changer. 
-With the first part of this paper, the structure and ideas of Che are explained in more detail, the second part of this pa-
-per covers general arguments about this approach and gives a detailed overview of reachable benefits for education.
+With the first part of this paper, the structure and ideas of Che are explained in more detail, the second part of this paper covers general arguments about this approach and gives a detailed overview of reachable benefits for education.
 
@@
 For Openshift the standard-openshift authentification is used.
+
 Eclipse Che  can be run  on any  base-technology  on  localhost – this is rather for demonstrations aswell as for developers working on the Che-Code. 
@@
 Eclipse Che does not work like a normal IDE. 
-Che consists of three main components shown in Figure \ref{fig:cheworkflow}\footnote{source: \url{https://en.wikipedia.org/wiki/Eclipse_Che#/media/File:Eclipse_Che_-_Workflow.PNG}}.
+Che consists of three main components shown in Figure \ref{fig:cheworkflow}\footnote{source: \url{https://en.wikipedia.org/wiki/Eclipse_Che\#/media/File:Eclipse_Che_-_Workflow.PNG}}.
 
@@
 \end{figure}
+\newpage
 
@@
 Having these runtimes inside a docker-container comes with great benefits, outlined later in this paper.
+
 The second component is the web-based IDE, which is basically a webpage performing the required REST-calls and utility-tasks for the developer. 
@@
 When starting a Maven-project, the workspace is initialized with Maven and Java.
+
 The last component is the Che-administration server. 
@@
 I expect this feature to be rather a proof of concept.
+
 A single workspace can contain multiple projects, whereof every project will have its own runtime. 
@@
 It's mostly dependent on the team which kind of cooperation they prefer.
+
 As  an  interesting  fact,  Eclipse  Che  is  developed  with Eclipse Che, proving that complex cloud development is possible. 
@@
 
+\subsection{Reproduction and Maintenance}
 The university of Arizona showed \cite{CO16}, that about 50\% of software inspected was not able to build. 
@@
 Experienced programmers are usually able to overcome these issues, given enough time. This continues to be an issue \cite{CL14}, and without a different approach probably always will be.
+
 This problem is solved, if the software is built inside a docker-container and follows the principles of self-containment and image-immutability \cite{BI17}.
@@
 Typically, after making the first iterations of your software on your local machine, dockering it will yield to 100 problems at once, not only being frustrating but taking tons of time.
+
 With Che, this kind of Big-Bang-Migration is not possible, as from the very first step you are working inside of a docker-container.
@@
 However, once the errors are resolved, every further development will work more agile by design.
+
 The artifacts (whether containers or pods) produced with Che are self-contained and immutable, therefore every result is reproducible. 
@@
 The behavior of this multipod is unified and testable, making a stable release of full environments possible.
+
 This is one of the primary goals of the DevOps approach \cite{JA16}. 
 The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{he Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of microservices \cite{IV18}, while being agile and reproducible.
+
+\subsection{Faster Setup}
 Onboarding new teammembers is made easy with Che. 
@@
 It is also notable, that the (technical) onboarding of developers and operators is streamlined into one single process.
+
 The combination of a quick setup and the usability of every enhancement makes Che especially interesting to recruit troubleshooters into your projects, which solve some of the very specialiced tasks.
@@
 As a further example, the workspaces could be shut down out of normal business-time. This attribute is crucial for a rented infrastructure, which is paid commonly on both workload and storage. 
+\subsection{Always On}
 The very first handicap coming with Che is the requirement for a good internet connection. 
 While your normal buro is already equipped with internet, one may want to work on train or there is simply a problem with the WiFi in your hotel.
-With the possibility to setup che on localhost, you can learn the basics and routines offline. 
-As the main benefits come in terms of teamwork, this is not a real solution.
+With the possibility to setup che on localhost, you can learn the basics and routines offline. As the main benefits come in terms of teamwork, this is not a real solution.
+\subsection{Knowledge Required}
 The primary problem with Che is the requirement for an already dense knowledge of cloud related technology.
@@
 It's therefore necessary to get everyone on a certain base-level to solve their own problems, otherwise the expert will be overloaded and the project stuck.
+
+\subsection{Different Workflow}
 Additionally, the missing default “build and run” is a paradigm-shift for many developers, which will take a while to get used to. 
@@
 
-TODO CITE CITE CITE
-The primary shift that’s required is not one of new tooling, as most developers already have the basic tooling they need. 
-Rather, the needed shift is one of philosophy. 
-TODO CITE CITE CITE END
+\begin{centering}
+	\textit{"The primary shift that’s required is not one of new tooling, as most developers already have the basic tooling they need.\newline 
+	Rather, the needed shift is one of philosophy."}
+\end{centering}
 
@@
 It just means additional effort and know-how.
+
 The last problem is the vulnerability to malevolent users. 
@@
 It is therefore important to not only restrict access to project resources, but also the double-check and guard the releases.
+
 Some security risks can also emerge from docker-images as well \cite{PE13} but are out of scope for this paper.
@@
 For simplicity I will only elaborate universities, to which I can best relate.
+\subsection{New motivational and intuitive Courses}
 Eclipse Che provides the possibility to learn primary cloud-technologies hands on and in a simple and supportive manner. 
@@
 The student will need to understand cloud-components aswell as team-work to succeed with Che.
+
 The produced artifacts can simply be shipped as containers, giving comfortable access to the professor. 
@@
 With a clear reason on the usefulness of the task, the visible progress and the direct impact of personal effort should lead to intrinsic motivation, which studies show to have a strong effect on education \cite{DECI93}.
+
 As a requirement, it's necessary to provide a meaningful and fun task. 
@@
 It's not dependent on the individual students.
+
 In general, this approach enables studies to be less about the subject, but instead to be about learning the subject. 
@@
 Another benefit for education is about fairness in IT-Education.
+
+\subsection{Equality and Accessibility}
 With access to a university Che workspace, every person can develop the same software in the same environment. 
@@
 While these technologies are great and important for many applications, the required technologies are exclusive due to  their  costs.
+
 This  hindrance  for  education  is  usually adressed  with  IOS-laboratories  and/or  cloud-resources paid for by university. 
@@
 Making  assignments  mandatory, providing extrinsic motivation \cite{DECI93} has usually a negative impact on the subjective experience.
+
 A different problem comes with machine-learning, which is usually a process of trial and error to some extend.
 Long waiting times disconnect the action from the result being not beneficial for the learning process \cite{HU95}\cite{DECI93}.
+
 One may argue that having this kind of anticipation is extra rewarding and therefore highly motivational.
@@
 It should therefore be minimized.
+
 With studies showing basics of good learning experiences \cite{FI05}\cite{TC88} and the impact of motivation \cite{HU95}\cite{DECI93}, it is mandatory for institutions to provide the best possible environment.
@@
 Most universities already provide free services, such as software-suites, e-libraries or cloud storage.
+
 Simply providing a technical correct solution (like laboratories) is not enough to provide a good environment.
@@
 With more and more tools making it easier, it's just a matter of time until cloud-tec becomes standard instead of bonus.
+
 With Eclipse Che many problems about teaching students these Cloud Skills and reducing the initial scope for learning, are addressed. 
-While the impacts on motivation and the actual learning results are only hypothetical, most of the technical aspects are solved by Che.  
+While the impacts on motivation and the actual learning results are only hypothetical, most of the technical requirements are solved by Che.  
 As Eclipse Che is cutting edge it whatsoever lacks many quality of life features, has some bugs, and the design is undefined to some extend.
+
 It somewhat is a chance: 
@@
 With Che there is another chance to “get back into the game” – and provide corporation-free education.
+
 While this sounds harsh towards companies, I think unbiased education for students is not a nice thing to have, but a core concept of education in general. 
@@
 %% Starten Sie "biber paper", um eine Biliographie zu erzeugen.
-\printbibliography
-
+%\printbibliography
+\bibliography{che}
 \end{document}

### 2019-05-06 che.tex
Added Subsection for reproducible Research

@@
 This is one of the primary goals of the DevOps approach \cite{JA16}. 
-The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{he Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of microservices \cite{IV18}, while being agile and reproducible.
+The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{he Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of micro-services \cite{IV18}, while being agile and reproducible.
+\subsection{Sharing Results}
+The topic \textit{docker for reproducible research} has already been promoted \cite{BO15} and successfully applied to complex topics \cite{ST16} such as machine learning. 
+As Clark et al.
+\cite{CL14} have used code of the several repositories provided from their university, and I expect most of these projects to be of scientific nature, 50\% of the research results where not reproducible.
 
+The idea to not only ship code, but ship the experiment as a whole makes reproduction much easier. 
+With Che (and using best practices for containers) the experiments done will be containerized and therefore reproducible. 
+Even more complex environments including multiple components can be made in the same paradigms forming a pod. 
+Such a pod could test applications working on an cassandra-cluster, fullstack web-applications or network-traffic-simulations.
+
+Che also offers a whole new way of sharing the results: 
+If a peer wants to inspect the code, he can be granted an account and get hands on the experiment without any setup. 
+In an further attempt, a common guest-account can be established for anyone interested with read-only rights. 
+This does not only help cooperation and understanding, but is also a good way to verify the results. 
+At the moment we rather blindly trust each other about the results and data presented - we accepted that we are not able to reproduce it anyway or atleast not without inhuman effort.
+This changes with reproducible experiments, which verify the data and can help peers reviewing something even more important - the experiment itself.
 \subsection{Faster Setup}
@@
 
+\subsection{Vulnerability}
 The last problem is the vulnerability to malevolent users. 

### 2019-05-09 che.tex
typos and minor

@@
 from cloud computing, cloud storage, database-as-a-service or rentable authentication. 
-Modern Software needs to be scalable, self-contained, and should run on any device, aswell as in the cloud.  
+Modern Software needs to be scalable, self-contained, and should run on any device, as well as in the cloud.  
 This trend is also visible in Googles searches \ref{fig:digdocker}.
@@
 Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
-Che is unlike a normal eclipse distribution: 
+Che is unlike a normal Eclipse-distribution: 
 It is hosted on a server and developers get access with their browser. 
@@
 
-Eclipse Che  can be run  on any  base-technology  on  localhost – this is rather for demonstrations aswell as for developers working on the Che-Code. 
+Eclipse Che  can be run  on any  base-technology  on  localhost – this is rather for demonstrations as well as for developers working on the Che-Code. 
 When run (locally) on docker, the plugin CheDir enables a portable workspace, 
@@
 \subsection{Reproduction and Maintenance}
-The university of Arizona showed \cite{CO16}, that about 50\% of software inspected was not able to build. 
-Common reasons are missing, changed or faulty dependencies, leading to the common term of \textit{dependency hell} aswell as unclear documentation. 
+The University of Arizona showed \cite{CO16}, that about 50\% of software inspected was not able to build. 
+Common reasons are missing, changed or faulty dependencies, leading to the common term of \textit{dependency hell} as well as unclear documentation. 
 Experienced programmers are usually able to overcome these issues, given enough time. This continues to be an issue \cite{CL14}, and without a different approach probably always will be.
@@
 The very first handicap coming with Che is the requirement for a good internet connection. 
-While your normal buro is already equipped with internet, one may want to work on train or there is simply a problem with the WiFi in your hotel.
-With the possibility to setup che on localhost, you can learn the basics and routines offline. As the main benefits come in terms of teamwork, this is not a real solution.
+While a common office is already equipped with internet, one may want to work on a train or there is simply a problem with the WiFi in the hotel.
+With the possibility to setup Che on localhost, you can learn the basics and routines offline. As the main benefits come in terms of teamwork, this is not a real solution.
 \subsection{Knowledge Required}
@@
 Even complicated topics, like microservices and IoT-applications can easily be adaptet inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
-The student will need to understand cloud-components aswell as team-work to succeed with Che.
+The student will need to understand cloud-components as well as team-work to succeed with Che.
 
@@
 One may argue that having this kind of anticipation is extra rewarding and therefore highly motivational.
-From my personal experience, working in this modus becomes exhausting after a period of weeks, especially when errors occur. 
-I expect this experience to be highly personal. 
-I think providing the fastest infrastructure possible, enabling the fastest results, is the better way. 
+From experience, working in this mode becomes exhausting after a period of weeks, especially when errors occur.
+While this experience is highly personal,providing the 
+fastest infrastructure possible, enabling the fastest results,is the better way.
 Making  progress  slower  is much easier than making it faster. 

### 2019-05-09 che.tex
Reduced Phrases and rewritten a lot

@@
 Cloud technologies are already state of the art, enabling global players like Google to run applications on a world-scale.
-The market-leaders got their portfolio stocked up with  a  variety  of  different  cloud  technologies,  ranging 
-from cloud computing, cloud storage, database-as-a-service or rentable authentication. 
-Modern Software needs to be scalable, self-contained, and should run on any device, as well as in the cloud.  
-This trend is also visible in Googles searches \ref{fig:digdocker}.
+The market-leaders stocked up their portfolios with  a  variety  of  different  cloud  technologies,  ranging from cloud computing, cloud storage, database-as-a-service or rentable authentication. 
+Modern software needs to be scalable, self-contained, and should run on any device, as well as in the cloud.  
+This trend is also visible in Google searches \ref{fig:digdocker}.
 
@@
 On-demand self-service, broad network-access, resource pooling, rapid elasticity and measured service \cite{ME11}. 
-Cloud technology usually splits into two components:  Infrastructure  and  Services,  where  docker and Kubernetes are considered infrastructure and provide a  platform  to  host  services.  
+Cloud technology usually splits into two components:  infrastructure  and  services,  where  Docker and Kubernetes are considered infrastructure and provide a  platform  to  host  services.  
 Most  of  the  attributes  are (pseudo-) achievable by the infrastructure alone – such as restarting containers. 
@@
 Lately the term “Cloud-native” came up \cite{BI17}, referring to software which is built from the very first steps to run inside the cloud.
-This kind of software is usually running inside a docker-container and both resilient and elasticby design \cite{TO17}.
+This kind of software is usually running inside a docker-container and both resilient and elastic by design \cite{TO17}.
 Resilience refers to the ability of handling failure, both from external sources and restarting/configuring themselves.  
 Elasticy is the ability to request more resources on demand, performing more computations when required, but also to free the resources if not needed. 
-Cloud-ready Software usually achieves only a certain degree of resilience.   
+Cloud-ready software usually achieves only a certain degree of resilience.   
 Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities \cite{MIN18}. 
@@
 
-So why are there no courses? There are many reasons, but the most prominent is the complexity of the topic: 
+SThere are many reasons why there are not many courses, but the most prominent is the complexity of the topic: 
 To build a real cloud native application, the developer must understand virtualization, system-administration, development, 
@@
 This is not to be mixed up with having multiple users work on the same project but on their own workspaces. 
-The code, and the workspace itself, can commonly be shared with any versioning tool like Git.
+The code, and the workspace itself, can commonly be shared with any versioning tool such as Git.
 It's mostly dependent on the team which kind of cooperation they prefer.
@@
 In addition to the previous mentioned benefits, there are more points about Che that are especially important for educational purposes, such as courses in the university, professional school or research departments.
-For simplicity I will only elaborate universities, to which I can best relate.
+For simplicity only universities will be addressed.
 \subsection{New motivational and intuitive Courses}
@@
 With fitting material provided, it's easy to learn cloud-native development, DevOps or Cloud-Operations. 
-Even complicated topics, like microservices and IoT-applications can easily be adaptet inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
+Even complicated topics, such as microservices and IoT-applications can easily be adaptet inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
 The student will need to understand cloud-components as well as team-work to succeed with Che.
@@
 With both the product and the code inside the workspace easily accessible, the professor can grant faster and better feedback for the students, which can have massive impact on learning \cite{HU95}\cite{HU18}.
-Additionally, the professor can provide automated end-to-end tests with tools like Jenkins or TravisCI. 
+Additionally, the professor can provide automated end-to-end tests with tools such as Jenkins or TravisCI. 
 Given such an environment, the student will immideatly get objective feedback, whether their code fulfills the requirements. 
@@
 According to PISA, this rate has not significally changed since the studies started in 2006. 
-Germany scored on the impact of socioeconomic-factors on education better than it did in 2006 but is worse than global average \cite{OECDCN15}.
+Germany scored better on the impact of socioeconomic-factors on education in 2015 than it did in 2006 but is worse than the global average \cite{OECDCN15}.
 The Mac-Laboratories and rented cloudspace are technically a good attempt at inclusion. 
@@
 It somewhat is a chance: 
-Instances like a university can have massive impact on the design of Che and drive the project to some extend for their own needs. 
+Instances such as a university can have massive impact on the design of Che and drive the project to some extend for their own needs. 
 With faculties working on the open source project, either providing code and features, or just testing Che in educational environments, both quality issues and missing features can be solved.    

### 2019-05-09 che.tex
removed me, myself and I

@@
 \label{sec:Introduction}
-As children we, used to look into the clouds and count sheep. 
-As developers, we look into the clouds and check our applications.  
+Children look into the clouds and count sheep. 
+Developers look into the clouds and check their applications.  
 Cloud technologies are already state of the art, enabling global players like Google to run applications on a world-scale.
@@
 This  will grant you access to autocompletion and other features, but you will miss Che specific helpers the web-IDE offers. 
-My personal oppinion is, that mounting remote file systems for development sounds like a infinite source of problems. 
-I expect this feature to be rather a proof of concept.
+Mounting remote file systems for development sounds lie an infinite source of problems.
+This feature looks to be rather a proof of concept.
 
@@
 As Clark et al.
-\cite{CL14} have used code of the several repositories provided from their university, and I expect most of these projects to be of scientific nature, 50\% of the research results where not reproducible.
+\cite{CL14} have used code of several repositories provided from their university, most of the projects can be expected scientific or experimental, 50\% of the research results were not reproducible.
 
@@
 In general, this approach enables studies to be less about the subject, but instead to be about learning the subject. 
-I consider this the primary goal of university courses. 
+This can be considered one of the primary goals of university courses. 
 Another benefit for education is about fairness in IT-Education.
@@
 
-While this sounds harsh towards companies, I think unbiased education for students is not a nice thing to have, but a core concept of education in general. 
+While this sounds harsh towards companies, unbiased education for students is not a nice thing to have, but a core concept of education in general. 
 Only if that is true, the students can make unbiased decisions and make unbiased research. 

### 2019-05-09 che.tex
removed You, We, His

@@
 remote environment. 
-Instead of developing on your local machine, a remote development-server is accessed via a webbrowser 
+Instead of developing on a local machine, a remote development-server is accessed via a webbrowser 
 and the artifacts are compiled into docker-containers.
@@
 \label{subsec:Tec}
-A normal IDE is software on your computer which makes programming easier than doing it in a plain text-editor and commandline. 
+A normal IDE is software on a computer which makes programming easier than doing it in a plain text-editor and commandline. 
 Common tasks for an IDE are dependency management, debugging, refactoring and autocompletion.
-It is important that an IDE is not required to develop – you can write .java-files in Vim, compile on the console and run a .Jar manually. 
-With that in mind, we can see that the IDE is just a tool to make things (a lot) more comfortable – but you also have requirements such as the JDK, the .Jars you need in your classpath etc.
-as they are only utilized by your IDE.
-You can consider this stack as your workspace as it contains everything required to work, with the IDE making the workspace comfortable to use.
+It is important that an IDE is not \textit{required} to develop - Vim can write Java files, the JDK can compile and the JRE can run everything manually from console. 
+With that in mind,an IDE is just a toolstack to make things (a lot) more comfortable – but it also has requirements such as the JDK, the .Jars needed in the classpath etc.
+as they are only utilized by the IDE.
+This stack can be considered a \textit{workspace} as it contains everything required to work, with the IDE making the workspace comfortable to use.
 Eclipse Che does not work like a normal IDE. 
@@
 \end{enumerate}
-A workspace is like your above mentioned common \textit{workspace}. 
+A workspace is like the above mentioned common \textit{workspace}. 
 It is a single machine, as a container, containing the project files, compilers, package managers and an IDE-Interface. 
@@
 \end{itemize}
-An additional distinction is about building your software.
-Instead of building the .Jar in your remote-workspace, it is build inside a container in the remote-workspace.
+An additional distinction is about building software.
+Instead of building the .Jar in the remote-workspace, it is build inside a container in the remote-workspace.
 While this sounds a bit confusing, it makes sense to separate the actual runtime from the workspace. 
-This is also what happens on your machine – you compile an executable and run it separately. 
+This is also what happens on a local machine – the executable is compiled and run separately. 
 Having these runtimes inside a docker-container comes with great benefits, outlined later in this paper.
@@
 The second component is the web-based IDE, which is basically a webpage performing the required REST-calls and utility-tasks for the developer. 
-It is sent by the administrative server when an authorized user accesses his workspace via browser. 
+It is sent by the administrative server when an authorized user accesses the workspace via browser. 
 These two components can be enriched with plugins.  
@@
 One in the workspace-API and one in the hosted web-IDE.  
-Che offers a suite of common plugins dependent on your project. 
+Che offers a suite of common plugins dependent on the project. 
 When starting a Maven-project, the workspace is initialized with Maven and Java.
@@
 The runtimes do not necessarily need to be on the same machines as their workspaces. 
-Instead of the web-based IDE, you can connect a desktop IDE to your workspace. 
-This is done by mounting the workspaces-filesystem  in  your  desktop  IDE. 
-This  will grant you access to autocompletion and other features, but you will miss Che specific helpers the web-IDE offers. 
+Instead of the web-based IDE, a desktop IDE can be connected to the workspace. 
+This is done by mounting the workspaces-filesystem in the desktop  IDE. 
+This  will grant access to autocompletion and other features, but will miss Che specific helpers the web-IDE offers. 
 Mounting remote file systems for development sounds lie an infinite source of problems.
@@
 The main challenge is to migrate existing software and fulfill these properties. 
-Typically, after making the first iterations of your software on your local machine, dockering it will yield to 100 problems at once, not only being frustrating but taking tons of time.
+Typically, after making the first iterations of the software on a local machine, dockering it will yield to 100 problems at once, not only being frustrating but taking tons of time.
 
-With Che, this kind of Big-Bang-Migration is not possible, as from the very first step you are working inside of a docker-container.
+With Che, this kind of Big-Bang-Migration is not possible, as from the very first step its working inside of a docker-container.
 The properties above can be hold with every change to the image and checked in a normal review-process. 
@@
 This is one of the primary goals of the DevOps approach \cite{JA16}. 
-The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{he Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of micro-services \cite{IV18}, while being agile and reproducible.
+The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{the Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of micro-services \cite{IV18}, while being agile and reproducible.
 \subsection{Sharing Results}
@@
 Che also offers a whole new way of sharing the results: 
-If a peer wants to inspect the code, he can be granted an account and get hands on the experiment without any setup. 
+If peers wants to inspect the code, they can be granted accounts and get hands on the experiment without any setup. 
 In an further attempt, a common guest-account can be established for anyone interested with read-only rights. 
 This does not only help cooperation and understanding, but is also a good way to verify the results.
-At the moment we rather blindly trust each other about the results and data presented - we accepted that we are not able to reproduce it anyway or atleast not without inhuman effort.
-This changes with reproducible experiments, which verify the data and can help peers reviewing something even more important - the experiment itself.
 \subsection{Faster Setup}
 Onboarding new teammembers is made easy with Che. 
-A newly joining member does not need to install IDEs, languages, tools and packages, he only needs access to his Che-Account and a singular workspace with every dependency build for him.
+A newly joining member does not need to install IDEs, languages, tools and packages, they only needs access to their Che-Account and a singular workspace with every dependency build for them.
 Depending on the complexitiy of the project this required atmost an hour. 
@@
 
-The combination of a quick setup and the usability of every enhancement makes Che especially interesting to recruit troubleshooters into your projects, which solve some of the very specialiced tasks.
+The combination of a quick setup and the usability of every enhancement makes Che especially interesting to recruit troubleshooters into projects, which solve some of the very specialiced tasks.  
 With  the  artifacts  being  reproducible,  the  expert  can quickly inspect the subject and analyse problems.
-With a finished workspace he can quickly start working. 
-With his changes being also in docker, they will be visible and functional for the whole team.
+With a finished workspace they can quickly start working. 
+With their changes being also in docker, they will be visible and functional for the whole team.
 Therefore,  the  hardest  common  problems  regarding \textit{quick help} are addressed and solved. 
@@
 While a common office is already equipped with internet, one may want to work on a train or there is simply a problem with the WiFi in the hotel.
-With the possibility to setup Che on localhost, you can learn the basics and routines offline. As the main benefits come in terms of teamwork, this is not a real solution.
+With the possibility to setup Che on localhost, the basics and routines can be learned offline. As the main benefits come in terms of teamwork, this is not a real solution.
 \subsection{Knowledge Required}
@@
 
-Intuitively this is true: Every developer needs to setup his own workspace – the skills to do this are already there. 
+Intuitively this is true: Every developer needs to setup their own workspace – the skills to do this are already there. 
 The only difference is to setup the workspace in a common, documentated and descriptive way.
@@
 Making  progress  slower  is much easier than making it faster. 
-In addition, the time spent waiting on your machine learning algorithm or moving to the mac-laboratory on weekends is simply dead time. 
+In addition, the time spent waiting on a machine learning algorithm or moving to the mac-laboratory on weekends is simply dead time. 
 It should therefore be minimized.
@@
 Simply providing a technical correct solution (like laboratories) is not enough to provide a good environment.
-One student with a better socio-economic background who can afford a 300\$ GPU will have a better baseline for his education. 
+One student with a better socio-economic background who can afford a 300\$ GPU will have a better baseline for their education. 
 This is simply put unfair.

### 2019-05-09 che.tex
feedback from tex spellchecker

@@
 \begin{abstract}
-Eclipse Che is a revolutionary cloud-based IDE, inheriting the approach of building cloud-ready software within a 
+Eclipse Che is a revolutionary cloud-based IDE, inheriting the approach of building cloud native software within a 
 remote environment. 
@@
 	\centering
-	\includegraphics[width=0.5\linewidth]{imgs/DigDocker}
+	\includegraphics[width=0.75\linewidth]{imgs/DigDocker}
 	\caption{Rising interest in Cloud Tec, Data provided by Google Trends}
@@
 Figure \ref{fig:digdocker} shows the steadily rising interest in the most popular cloud technologies. 
-Docker is a container-engine enabling OS-Virtualization, Kubernetes is a Container-Platform manager and Openshift is a software-suite based on Kubernetes. 
+Docker is a container-engine enabling OS-Virtualization, Kubernetes is a Container-Platform manager and OpenShift is a software-suite based on Kubernetes. 
 These technologies will be summarized in \ref{subsec:Req}.
@@
 “Cloud-Ready” – meaning that it can be run in common cloud environments, such as Kubernetes. 
-The term cloudready  does  usually  not  include  live-scalability.
-Cloud-readieness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a dockercontainer.
+The term cloud ready  does  usually  not  include  live-scalability.
+Cloud-readieness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a docker-container.
 Lately the term “Cloud-native” came up \cite{BI17}, referring to software which is built from the very first steps to run inside the cloud.
@@
 The applications must also be able to use the gained resources in a productive and healthy way.  
-Modern  cloud-platforms  can  perform  pseudo-scalabilty for cloud-ready-services by starting multiple instances of the same service. 
+Modern  cloud-platforms  can  perform  pseudo-scalability for cloud ready-services by starting multiple instances of the same service. 
 The cloud-platform performs load-balancing between these instances. 
 This naive approach can work out fine, but for example starting two services working on the same database will not yield any real scalability. 
-Another common problem arises when two services are started on the same machine and need the same port, just to name two problems with cloud-ready  software in a cloud environment.
+Another common problem arises when two services are started on the same machine and need the same port, just to name two problems with cloud ready  software in a cloud environment.
 
@@
 universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
-There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud-native development is sealed behind corporate doors and conference workshops.
-Meanwhile,  the  annual  Stackoverflow  developersurvey \cite{SO19} showed, that DevOps-engineers are not only 
-amongst the most wanted and most paid jobs but also tend to be the happiest participants.  
+There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud native development is sealed behind corporate doors and conference workshops.
+Meanwhile,  the  annual  StackOverflow  developer-survey \cite{SO19} showed, that DevOps-engineers are not only 
+among the most wanted and best paid jobs but also tend to be the happiest participants.  
 Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career.
 
-SThere are many reasons why there are not many courses, but the most prominent is the complexity of the topic: 
+There are many reasons why there are not many courses, but the most prominent is the complexity of the topic: 
 To build a real cloud native application, the developer must understand virtualization, system-administration, development, 
@@
 Additionally, the developer needs to use specific tools, know best practices, tests and work in a team.
-All these factors make a high stake for entrance, also  represented in the developersurvey \cite{SO19} which states that most DevOps engineers have a decade of experience, usually in operations topics.
+All these factors make a high stake for entrance, also  represented in the developer-survey \cite{SO19} which states that most DevOps engineers have a decade of experience, usually in operations topics.
 
-Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
+Luckily a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud native development.
 Che is unlike a normal Eclipse-distribution: 
 It is hosted on a server and developers get access with their browser. 
-Instead of installing dependencies, compilers etc. locally, a common workspace is set up, which is shared amongst the programmers. 
-The software build is run inside a docker-container, making it cloud-native by default. 
+Instead of installing dependencies, compilers etc. locally, a common workspace is set up, which is shared among the programmers. 
+The software build is run inside a docker-container, making it cloud native by default. 
 With Version 7 upcoming and prominent support from RedHat, Eclipse Che is not a prototype anymore. 
@@
 \label{subsec:Req}
-Che is hosted as a server-application and is already cloudnative. 
+Che is hosted as a server-application and is already cloud native. 
 The only requirements are that either Docker, Kubernetes or OpenShift are available for the installation.
@@
 
-Docker\footnote{https://www.docker.com/} is a container-plattform, where a container is a standardized unit of of software including the code, dependencies and core-functionalities. 
+Docker\footnote{https://www.docker.com/} is a container-platform, where a container is a standardized unit of of software including the code, dependencies and core-functionalities. 
 A container runs on an OS-Virtualization, originating from LXC, has declared interfaces and can be parameterized. 
 One of the core features is to connect containers for bigger projects, e.g. one 
-database container and one web-server container, which are connected into a virtual network. 
+database container and one web-server-container, which are connected into a virtual network. 
 Docker is mostly famous for this docker-engine, but enriches it with monitoring and logging, as well as the ability to move containers onto different machines.
@@
 Kubernetes\footnote{https://kubernetes.io/} is a container-orchestration-system based on Docker. 
-Kubernetes picked up the growing problem in managing multiple docker-containers by  simplyfing  resource-management. 
+Kubernetes picked up the growing problem in managing multiple docker-containers by  simplifying  resource-management. 
 Additional to deployment and monitoring of container-groups it provides auto-restart mechanisms and scaling based on (custom) metrics. 
@@
 The term pod, also later used in this paper, refers to a suite of connected containers. 
-An example pod would be a simple two-tier web-application, where the database and webserver are two different, but connected containers.
+An example pod would be a simple two-tier web-application, where the database and web-server are two different, but connected containers.
 
-Openshift\footnote{https://www.openshift.com/} provides a software-suite around Kubernetes with the goal to automate cloud-native development and delivery.
+OpenShift\footnote{https://www.openshift.com/} provides a software-suite around Kubernetes with the goal to automate cloud native development and delivery.
 Notable  additions  include  Jenkins,  Gitea,  Sonarqube and lately Che. 
 Regarding Che there are no notable changes in handling, as Che is simply run on the build in Kubernetes.
-When run on Kubernetes or Docker Che will require two additional containers for authentification with KeyCloack. 
-For Openshift the standard-openshift authentification is used.
+When run on Kubernetes or Docker Che will require two additional containers for authentication with KeyCloack. 
+For OpenShift the standard-openshift authentication is used.
 
@@
 A normal IDE is software on a computer which makes programming easier than doing it in a plain text-editor and commandline. 
-Common tasks for an IDE are dependency management, debugging, refactoring and autocompletion.
+Common tasks for an IDE are dependency management, debugging, refactoring and auto-completion.
 It is important that an IDE is not \textit{required} to develop - Vim can write Java files, the JDK can compile and the JRE can run everything manually from console. 
@@
 The authorization is also done at the admin-server and not by every workspace.
-The workspaces can be split amongst multiple machines.
-The runtimes do not necessarily need to be on the same machines as their workspaces. 
+The workspaces can be shipped onto multiple separate machines.
+The runtimes do not necessarily need to be on the same machines as their workspace. 
 Instead of the web-based IDE, a desktop IDE can be connected to the workspace. 
 This is done by mounting the workspaces-filesystem in the desktop  IDE. 
-This  will grant access to autocompletion and other features, but will miss Che specific helpers the web-IDE offers. 
+This  will grant access to auto-completion and other features, but will miss Che specific helpers the web-IDE offers. 
 Mounting remote file systems for development sounds lie an infinite source of problems.
@@
 Projects can either share a versioning control or be separated. 
-You can build whole pods in a single workspace with the single containers being the subprojects. 
-You can also compose applications inside the workspaces, meaning that the resulting 
+You can build whole pods in a single workspace with the single containers being the sub-projects. 
+You can also compose applications inside the workspace, meaning that the resulting 
 container of a project can be the base-image of another project. 
-A workspace can be shared amongst multiple users. 
+A workspace can be shared among multiple users. 
 Both will work on the same remote filesystem, altering files simultaneous. 
@@
 drive-documents. 
-This is not to be mixed up with having multiple users work on the same project but on their own workspaces. 
+This is not to be mixed up with having multiple users work on the same project but on their own workspace. 
 The code, and the workspace itself, can commonly be shared with any versioning tool such as Git.
@@
 First it reduces failure when migrating software into the cloud and enhances the experience of working in a rich remote environment.
-Second is to set up people into existing projects faster and failsafe.
+Second is to set up people into existing projects faster and fail safe.
 
@@
 The self-containment principle dictates that  at build-time every required dependency is given, and a fully functional image is created.
-A dockerized maven-application would include java, maven and every required maven package to run the application in the current state. 
+A containerized maven-application would include java, maven and every required maven package to run the application in the current state. 
 The image-immutability means to forbid alternations to the environment. 
@@
 The main challenge is to migrate existing software and fulfill these properties. 
-Typically, after making the first iterations of the software on a local machine, dockering it will yield to 100 problems at once, not only being frustrating but taking tons of time.
+Typically, after making the first iterations of the software on a local machine, containerization will yield to 100 problems at once, not only being frustrating but taking tons of time.
 
-With Che, this kind of Big-Bang-Migration is not possible, as from the very first step its working inside of a docker-container.
+With Che, this kind of Big-Bang-Migration is not possible, as from the very first step it is working inside of a docker-container.
 The properties above can be hold with every change to the image and checked in a normal review-process. 
@@
 You can keep multiple projects in a single workspace and launch multiple pods at once, running the full environment. 
-The behavior of this multipod is unified and testable, making a stable release of full environments possible.
+The behavior of this multi-pod is unified and testable, making a stable release of full environments possible.
 
 This is one of the primary goals of the DevOps approach \cite{JA16}. 
-The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{the Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of micro-services \cite{IV18}, while being agile and reproducible.
+The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{the Plugin Arktik for Eclipse offered IoT Development with Che.
+While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of micro-services \cite{IV18}, while being agile and reproducible.
 \subsection{Sharing Results}
@@
 A newly joining member does not need to install IDEs, languages, tools and packages, they only needs access to their Che-Account and a singular workspace with every dependency build for them.
-Depending on the complexitiy of the project this required atmost an hour. 
+Depending on the complexity of the project this required at most an hour. 
 While this may seem long, the setup time is dependent on the complexity of the workspace. 
@@
 The work for this is done automatic and remote, so the user is free to dive into the documentation or have some coffee. 
-His workspace will be able to launch the latest version immidieatly.
+His workspace will be able to launch the latest version immediately.
 It is also notable, that the (technical) onboarding of developers and operators is streamlined into one single process.
 
-The combination of a quick setup and the usability of every enhancement makes Che especially interesting to recruit troubleshooters into projects, which solve some of the very specialiced tasks.  
-With  the  artifacts  being  reproducible,  the  expert  can quickly inspect the subject and analyse problems.
+The combination of a quick setup and the usability of every enhancement makes Che especially interesting to recruit troubleshooters into projects, which solve some of the very specialized tasks.  
+With  the  artifacts  being  reproducible,  the  expert  can quickly inspect the subject and analyze problems.
 With a finished workspace they can quickly start working. 
@@
 The workspaces collect metrics like any other container and are therefore scalable on demand. 
-The ability of downscaling the development resources has a big impact on the overall 
+The ability of down-scaling the development resources has a big impact on the overall 
 workload. 
-As a further example, the workspaces could be shut down out of normal business-time. This attribute is crucial for a rented infrastructure, which is paid commonly on both workload and storage. 
+As a further example, the workspace could be shut down out of normal business-time. This attribute is crucial for a rented infrastructure, which is paid usually on both workload and storage. 
 \subsection{Always On}
@@
 Setting up the Workspace, docker-images, build and CI scripts is a complex topic and can not be done by someone getting started. 
-This requirement extends when working with a multipod-environment. 
+This requirement extends when working with a multi-pod-environment. 
 While most of the tasks are by themselves rather simple, the bare amount of such tasks (especially when migrating an existing project to Che) requires a broad knowledge.
 This can be done by a single expert user, enabling the other developers  to  use  the  build-routines and  environments provided. 
-However, many problems which arise are related to the cloud-environement and will be hard to solve by a novice user, making him dependent on the expert. 
+However, many problems which arise are related to the cloud-environment and will be hard to solve by a novice user, making him dependent on the expert. 
 It's therefore necessary to get everyone on a certain base-level to solve their own problems, otherwise the expert will be overloaded and the project stuck.
@@
 Intuitively this is true: Every developer needs to setup their own workspace – the skills to do this are already there. 
-The only difference is to setup the workspace in a common, documentated and descriptive way.
+The only difference is to setup the workspace in a common, documented and descriptive way.
 
 Another problem with Che is the uncommon debugging. 
-As the code is running in a remote-runtime, the common breakpoint  and  valuecheck
-is  missing.  
+As the code is running in a remote-runtime, the common breakpoints  and  value checks are missing.  
 Instead,  one  either must be familiar with the languages commandline-debugging or write good logs and metrics. 
@@
 The debugging and error-analysis inside the container is the same as it would be in the productive environment. 
-Having meaningful logs with different loglevels and collecting valuable metrics is 
-in general a good attribute for software in any circumstance. 
+Having meaningful logs with different loglevels and collecting valuable metrics is in general a good attribute for software in any circumstance. 
 It just means additional effort and know-how.
@@
 This  can  also  be achieved by human error. 
-Dependent on the setup, the attacker could also auto-deploy hidden spyware into a pro-
-duction environment. 
+Dependent on the setup, the attacker could also auto-deploy hidden spyware into a production environment. 
 While this is also a problem for normal developer teams, the fast pace of deployments with Che requires additional care. 
@@
 Eclipse Che provides the possibility to learn primary cloud-technologies hands on and in a simple and supportive manner. 
-Unfortunetally, these topics are currently not widely teached at university.
+Unfortunately, these topics are currently not widely teached at university.
 Che makes it possible to teach students fundamentals of modern software development.
 Depending on the courses subject, one can either provide scalable software or a running environment. 
-With fitting material provided, it's easy to learn cloud-native development, DevOps or Cloud-Operations. 
+With fitting material provided, it's easy to learn cloud native development, DevOps or Cloud-Operations. 
 Even complicated topics, such as microservices and IoT-applications can easily be adaptet inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
@@
 Additionally, the professor can provide automated end-to-end tests with tools such as Jenkins or TravisCI. 
-Given such an environment, the student will immideatly get objective feedback, whether their code fulfills the requirements. 
+Given such an environment, the student will immediately get objective feedback, whether their code fulfills the requirements. 
 This immediate feedback is a big motivator and has a positive impact on the learning experience \cite{FI05}. 
@@
 With a single investment for the tests, feedback for every student is granted at every time. 
-With the single investment of providing a running Che instance (and a short lecture how to use it), feedback on the artifacts and code can be immidieatly granted to every student, without having the overhead of setting up workspaces.
+With the single investment of providing a running Che instance (and a short lecture how to use it), feedback on the artifacts and code can be immediately granted to every student, without having the overhead of setting up the workspace.
 These investments can be done before the course starts at any time and can be reused.
@@
 With access to a university Che workspace, every person can develop the same software in the same environment. 
-A wellknown problem is the necessity of an IOS-build-device for IOS-appdevelopment.
+A well known problem is the necessity of an IOS-build-device for IOS-appdevelopment.
 Students which are limited in their access to the university-laboratory or cannot afford a mac are therefore handicapped.
@@
 Depending on the task, training neural networks can be 20 times faster if a CUDA-compatible graphics card is used. 
-The range for these graphiccards start at 300\$ and are therefore not commodity hardware,especially for students.  
+The range for these graphic-cards start at 300\$ and are therefore not commodity hardware,especially for students.  
 While these technologies are great and important for many applications, the required technologies are exclusive due to  their  costs.
 
-This  hindrance  for  education  is  usually adressed  with  IOS-laboratories  and/or  cloud-resources paid for by university. 
+This  hindrance  for  education  is  usually addressed  with  IOS-laboratories  and/or  cloud-resources paid for by university. 
 The topic of socio-economic factors is beyond the scope of this work, but the impact has been analyzed by the PISA Studies \cite{PISA15}, which state that socio-economicly handicapped students fail three times more likely to achieve a basic level of performance.
-According to PISA, this rate has not significally changed since the studies started in 2006. 
+According to PISA, this rate has not significantly changed since the studies started in 2006. 
 Germany scored better on the impact of socioeconomic-factors on education in 2015 than it did in 2006 but is worse than the global average \cite{OECDCN15}.

### 2019-05-12 che.tex
Changed KeyWords,minor spelling

@@
 remote environment. 
-Instead of developing on a local machine, a remote development-server is accessed via a webbrowser 
+Instead of developing on a local machine, a remote development-server is accessed via web-browser 
 and the artifacts are compiled into docker-containers.
@@
 \begin{keywords}
-Eclipse Che \and Cloud-Ready \and Cloud-Native \and DevOps \and Docker \and Kubernetes \and OpenShift
+Eclipse Che \and Cloud Native \and DevOps \and Docker \and Kubernetes \and Higher Education \and Software Development
 \end{keywords}
@@
 Cloud technologies are already state of the art, enabling global players like Google to run applications on a world-scale.
-The market-leaders stocked up their portfolios with  a  variety  of  different  cloud  technologies,  ranging from cloud computing, cloud storage, database-as-a-service or rentable authentication. 
+The market-leaders stocked up their portfolios with  a  variety  of  different  cloud  technologies,  ranging from cloud computing, cloud storage, database-as-a-service or rented authentication. 
 Modern software needs to be scalable, self-contained, and should run on any device, as well as in the cloud.  
@@
 Software running as a service can be put into two categories: 
-“Cloud-Ready” – meaning that it can be run in common cloud environments, such as Kubernetes. 
+\textit{Cloud ready} – meaning that it can be run in common cloud environments, such as Kubernetes. 
 The term cloud ready  does  usually  not  include  live-scalability.
-Cloud-readieness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a docker-container.
-Lately the term “Cloud-native” came up \cite{BI17}, referring to software which is built from the very first steps to run inside the cloud.
+Cloud-readiness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a docker-container.
+Lately the term \textit{Cloud native} came up \cite{BI17}, referring to software which is built from the very first steps to run inside the cloud.
 This kind of software is usually running inside a docker-container and both resilient and elastic by design \cite{TO17}.
 Resilience refers to the ability of handling failure, both from external sources and restarting/configuring themselves.  
-Elasticy is the ability to request more resources on demand, performing more computations when required, but also to free the resources if not needed. 
+Elasticity is the ability to request more resources on demand, performing more computations when required, but also to free the resources if not needed. 
 Cloud-ready software usually achieves only a certain degree of resilience.   
@@
 universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
-There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud native development is sealed behind corporate doors and conference workshops.
+There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloud-platforms and cloud native development is sealed behind corporate doors and conference workshops.
 Meanwhile,  the  annual  StackOverflow  developer-survey \cite{SO19} showed, that DevOps-engineers are not only 
@@
 It is important that an IDE is not \textit{required} to develop - Vim can write Java files, the JDK can compile and the JRE can run everything manually from console. 
-With that in mind,an IDE is just a toolstack to make things (a lot) more comfortable – but it also has requirements such as the JDK, the .Jars needed in the classpath etc.
+With that in mind,an IDE is just a tool-stack to make things (a lot) more comfortable – but it also has requirements such as the JDK, the .Jars needed in the classpath etc.
 as they are only utilized by the IDE.
@@
 
-The second component is the web-based IDE, which is basically a webpage performing the required REST-calls and utility-tasks for the developer. 
+The second component is the web-based IDE, which is basically a web-page performing the required REST-calls and utility-tasks for the developer. 
 It is sent by the administrative server when an authorized user accesses the workspace via browser. 
@@
 The last component is the Che-administration server. 
-In the administration-server the workspaces are orchestrated and monitored. 
+In the administration-server workspaces are orchestrated and monitored. 
 The authorization is also done at the admin-server and not by every workspace.
@@
 Even more complex environments including multiple components can be made in the same paradigms forming a pod. 
-Such a pod could test applications working on an cassandra-cluster, fullstack web-applications or network-traffic-simulations.
+Such a pod could test applications working on an Cassandra-cluster, fullstack web-applications or network-traffic-simulations.
 
@@
 With fitting material provided, it's easy to learn cloud native development, DevOps or Cloud-Operations. 
-Even complicated topics, such as microservices and IoT-applications can easily be adaptet inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
+Even complicated topics, such as micro-services and IoT-applications can easily be adapted inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
 The student will need to understand cloud-components as well as team-work to succeed with Che.
@@
 Germany scored better on the impact of socioeconomic-factors on education in 2015 than it did in 2006 but is worse than the global average \cite{OECDCN15}.
-The Mac-Laboratories and rented cloudspace are technically a good attempt at inclusion. 
+The Mac-Laboratories and rented cloud space are technically a good attempt at inclusion. 
 But students which are required to get to university to do their assignments are 

### 2019-06-12 che.tex
Minor removal and fixes

@@
 \begin{abstract}
-Eclipse Che is a revolutionary cloud-based IDE, inheriting the approach of building cloud native software within a 
+Eclipse Che is a open source cloud-based IDE, inheriting the approach of building cloud native software within a 
 remote environment. 
@@
 among the most wanted and best paid jobs but also tend to be the happiest participants.  
-Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career.
+Every student either wants to be happy or to be rich (some want both), making Cloud an interesting topic for their career.
 
@@
 This  will grant access to auto-completion and other features, but will miss Che specific helpers the web-IDE offers.
-Mounting remote file systems for development sounds lie an infinite source of problems.
 This feature looks to be rather a proof of concept.
@@
 For version 7 a multi-cursor file-editing will be provided, like it`s common in 
-drive-documents. 
+OneDrive-documents. 
 This is not to be mixed up with having multiple users work on the same project but on their own workspace. 
@@
 It's mostly dependent on the team which kind of cooperation they prefer.
-
-As  an  interesting  fact,  Eclipse  Che  is  developed  with Eclipse Che, proving that complex cloud development is possible. 
-This so-called \textit{dogfooding} usually leads to a product which puts requirements over any a-priori design or foreign interests.  
-Existing components will be enhanced if working with them becomes troublesome, leaving a clean and valuable core product.
 
@@
 
-\subsection{Reproduction and Maintenance}
+\subsection{Reproducibility and Maintenance}
 The University of Arizona showed \cite{CO16}, that about 50\% of software inspected was not able to build. 
@@
 As Clark et al.
-\cite{CL14} have used code of several repositories provided from their university, most of the projects can be expected scientific or experimental, 50\% of the research results were not reproducible.
+\cite{CL14} have used code of several repositories provided from their university, most of the projects can be expected scientific or experimental, therefore 50\% of the research results were not reproducible.
 
@@
 \subsection{Always On}
-The very first handicap coming with Che is the requirement for a good internet connection. 
+The very first drawback coming with Che is the requirement for a good internet connection. 
 While a common office is already equipped with internet, one may want to work on a train or there is simply a problem with the WiFi in the hotel.
@@
 Depending on the courses subject, one can either provide scalable software or a running environment. 
-With fitting material provided, it's easy to learn cloud native development, DevOps or Cloud-Operations. 
+With suitable material provided, it's easy to learn cloud native development, DevOps or Cloud-Operations. 
 Even complicated topics, such as micro-services and IoT-applications can easily be adapted inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 
@@
 
-The produced artifacts can simply be shipped as containers, giving comfortable access to the professor. 
-With both the product and the code inside the workspace easily accessible, the professor can grant faster and better feedback for the students, which can have massive impact on learning \cite{HU95}\cite{HU18}.
-Additionally, the professor can provide automated end-to-end tests with tools such as Jenkins or TravisCI. 
+The produced artifacts can simply be shipped as containers, giving comfortable access to the teaching staff. 
+With both the product and the code inside the workspace easily accessible, the teaching staff can grant faster and better feedback for the students, which can have massive impact on learning \cite{HU95}\cite{HU18}.
+Additionally, the teacher can provide automated end-to-end tests with tools such as Jenkins or TravisCI. 
 Given such an environment, the student will immediately get objective feedback, whether their code fulfills the requirements. 
@@
 While the impacts on motivation and the actual learning results are only hypothetical, most of the technical requirements are solved by Che.  
-As Eclipse Che is cutting edge it whatsoever lacks many quality of life features, has some bugs, and the design is undefined to some extend.
+As Eclipse Che is cutting edge it whatsoever lacks many quality of life features, has some bugs, and the preferred design is undefined to some extend.
 

### 2019-06-12 che.tex
Addition to future work

@@
 
-\documentclass[utf8]{lni}
+\documentclass[english,utf8]{lni}
 
@@
 This  can  also  be achieved by human error. 
-Dependent on the setup, the attacker could also auto-deploy hidden spyware into a production environment. 
-While this is also a problem for normal developer teams, the fast pace of deployments with Che requires additional care. 
-A rich CI/CD environment, such as a well-maintained Jenkins may lead to less-detailed code-reviews, as both unit- and end-to-end-tests succeed.  
 It is therefore important to not only restrict access to project resources, but also the double-check and guard the releases.
@@
 \label{sec:FutureWork}
-As this paper only summarizes benefits and outlines some ideas, the next step would be to put it to the test.
-In addition, a study could be elaborated on both the acceptance and effect of a CI/CD based development course with Che. 
-Unfortunately, the study would lack any comparison as there are few courses on this topic. 
+As this paper only argues about the possible improvement of education using Che, it should be therefore tested. 
+A study could be hold on both the acceptance and effect of a CI/CD based development course with Che in comparison to a conventional course. 
 These studies could be done with universities and corporations, which 
 may yield very different results.
+
+Additionally, it should be inspected whether the studies on general motivation and education hold for IT-students.
+
+When moving the scope from student to educator, a study can be hold about the time spent of the teacher on which part of the course and whether a positive effect on effective feedback-time can be achieved.
 

### 2019-06-12 che.tex
some removal

@@
 
-One may argue that having this kind of anticipation is extra rewarding and therefore highly motivational.
-From experience, working in this mode becomes exhausting after a period of weeks, especially when errors occur.
-While this experience is highly personal,providing the 
-fastest infrastructure possible, enabling the fastest results,is the better way.
-Making  progress  slower  is much easier than making it faster. 
-In addition, the time spent waiting on a machine learning algorithm or moving to the mac-laboratory on weekends is simply dead time. 
-It should therefore be minimized.
-
 With studies showing basics of good learning experiences \cite{FI05}\cite{TC88} and the impact of motivation \cite{HU95}\cite{DECI93}, it is mandatory for institutions to provide the best possible environment.

### 2019-06-21 che.tex
Added Formalia
added text to competitors and theia

@@
 %\startpage{-} % Beginn der Seitenzählung für diesen Beitrag
-%\editor{-}    % Namen der Herausgeber
-\booktitle{Under Revision for SKILL 2019} % Name des Tagungsband
+\editor{Gesellschaft für Informatik}    % Namen der Herausgeber
+\booktitle{SKILL 2019} % Name des Tagungsband
+%\band{Lecture Note in Informatics (LNI) – Seminars, Gesellschaft für Informatik, Bonn 2019}
 \year{2019}
@@
 Developers look into the clouds and check their applications.  
-Cloud technologies are already state of the art, enabling global players like Google to run applications on a world-scale.
+Cloud technologies are already state of the art, enabling global players to run applications on a world-scale.
 The market-leaders stocked up their portfolios with  a  variety  of  different  cloud  technologies,  ranging from cloud computing, cloud storage, database-as-a-service or rented authentication. 
@@
 which can be used on any docker based Che.
+
 \subsection{Technology and Workflow}
@@
 You can build whole pods in a single workspace with the single containers being the sub-projects. 
-You can also compose applications inside the workspace, meaning that the resulting 
-container of a project can be the base-image of another project. 
+You can also compose applications inside the workspace, meaning that the resulting container of a project can be the base-image of another project. 
 A workspace can be shared among multiple users. 
@@
 Due to changes being transmitted as HTTP-requests, the current implementation when working on the same file is a \textit{last write wins}-policy. 
-For version 7 a multi-cursor file-editing will be provided, like it`s common in 
-OneDrive-documents. 
+For version 7 a multi-cursor file-editing will be provided, like it`s common in OneDrive-documents. 
 This is not to be mixed up with having multiple users work on the same project but on their own workspace. 
@@
 It's mostly dependent on the team which kind of cooperation they prefer.
+
+Eclipse Che originates from the Eclise-Theia project. 
+Theia focussed (only) on the idea of having a remote workspace and access it via browser. 
+Che uses Theia and enriches it with idea of capsulation and containerization. 
+An alternative to Eclipse Che is Amazons Cloud9\footnote{https://aws.amazon.com/cloud9/}, which is free for students. 
+Microsoft is currently working on a similiar solution, the plattform Coder\footnote{https://coder.com/} is on a state similiar to Eclipse-Theia. 
+Che is the only completely free and open sourced Cloud-IDE.
 

### 2019-06-21 che.tex
Changed title

@@
 \begin{document}
-%%% Mehrere Autoren werden durch \and voneinander getrennt.
-%%% Die Fußnote enthält die Adresse sowie eine E-Mail-Adresse.
-%%% Das optionale Argument (sofern angegeben) wird für die Kopfzeile verwendet.
-\title[Eclipse Che for Education]{Eclipse Che for Fair and Easy Education}
-%%%\subtitle{Untertitel / Subtitle} % falls benötigt
+\title[theoretical evaluation of the potential advantages of cloud IDEs for research and didactic]{Theoretical evaluation of the potential advantages of cloud IDEs for research and didactic}
+\subtitle{on the example of Eclipse Che} % falls benötigt
 \author[Leonhard Applis]
@@
 It is therefore important to not only restrict access to project resources, but also the double-check and guard the releases.
-
-Some security risks can also emerge from docker-images as well \cite{PE13} but are out of scope for this paper.
-
 \section{Benefits for Education}
@@
 Eclipse Che provides the possibility to learn primary cloud-technologies hands on and in a simple and supportive manner. 
-Unfortunately, these topics are currently not widely teached at university.
 Che makes it possible to teach students fundamentals of modern software development.

### 2019-06-21 che.tex
new Title, slimmed

@@
 \begin{document}
-\title[theoretical evaluation of the potential advantages of cloud IDEs for research and didactic]{Theoretical evaluation of the potential advantages of cloud IDEs for research and didactic}
+\title[theoretical evaluation of the potential advantages of cloud IDEs for research and didactics]{Theoretical evaluation of the potential advantages of cloud IDEs for research and didactics}
 \subtitle{on the example of Eclipse Che} % falls benötigt
@@
 The behavior of this multi-pod is unified and testable, making a stable release of full environments possible. 
-
 This is one of the primary goals of the DevOps approach \cite{JA16}. 
-The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{the Plugin Arktik for Eclipse offered IoT Development with Che.
-While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of micro-services \cite{IV18}, while being agile and reproducible.
 \subsection{Sharing Results}
@@
 Even more complex environments including multiple components can be made in the same paradigms forming a pod.
-Such a pod could test applications working on an Cassandra-cluster, fullstack web-applications or network-traffic-simulations.
 
@@
 With  the  artifacts  being  reproducible,  the  expert  can quickly inspect the subject and analyze problems.
-With a finished workspace they can quickly start working. 
 With their changes being also in docker, they will be visible and functional for the whole team.
@@
 As a further example, the workspace could be shut down out of normal business-time. This attribute is crucial for a rented infrastructure, which is paid usually on both workload and storage. 
-\subsection{Always On}
-The very first drawback coming with Che is the requirement for a good internet connection. 
-While a common office is already equipped with internet, one may want to work on a train or there is simply a problem with the WiFi in the hotel.
-With the possibility to setup Che on localhost, the basics and routines can be learned offline. As the main benefits come in terms of teamwork, this is not a real solution.
 \subsection{Knowledge Required}
@@
 These investments can be done before the course starts at any time and can be reused.
-It's not dependent on the individual students.
-
-In general, this approach enables studies to be less about the subject, but instead to be about learning the subject. 
-This can be considered one of the primary goals of university courses. 
-Another benefit for education is about fairness in IT-Education.
-
 \subsection{Equality and Accessibility}
@@
 Depending on the task, training neural networks can be 20 times faster if a CUDA-compatible graphics card is used.
-The range for these graphic-cards start at 300\$ and are therefore not commodity hardware,especially for students.  
 While these technologies are great and important for many applications, the required technologies are exclusive due to  their  costs.

### 2019-06-21 che.tex
minor fixes, printable version

@@
 This can be done by a single expert user, enabling the other developers  to  use  the  build-routines and  environments provided.
-However, many problems which arise are related to the cloud-environment and will be hard to solve by a novice user, making him dependent on the expert. 
-It's therefore necessary to get everyone on a certain base-level to solve their own problems, otherwise the expert will be overloaded and the project stuck.
 

### 2019-06-21 che.tex
Manual changes, removed 2 weaker sources

@@
 Cloud-ready software usually achieves only a certain degree of resilience.   
-Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities \cite{MIN18}. 
+Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities. 
 The applications must also be able to use the gained resources in a productive and healthy way.  
@@
 Problems will arise one at a time and be solved one at a time, whether they are about code or operations. 
-Che therefore embraces the core-ideas of DevOps \cite{HUE12}\cite{JA16} and agile software development \cite{EBE16}.
+Che therefore embraces the core-ideas of DevOps \cite{JA16} and agile software development \cite{EBE16}.
 That being said, of course it's possible to big-bang-migrate into Che, causing the normal problem mentioned above. 
@@
 This is one of the primary goals of the DevOps approach \cite{JA16}. 
+\newpage
 \subsection{Sharing Results}
@@
 With Che, universities can provide a free, good and accessible development environment for every student.
+\newpage
 \section{Conclusion}

### 2019-07-13 che.tex
removed even more, so i only have 12 pages

@@
 Modern software needs to be scalable, self-contained, and should run on any device, as well as in the cloud.
-This trend is also visible in Google searches \ref{fig:digdocker}.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.75\linewidth]{imgs/DigDocker}
-	\caption{Rising interest in Cloud Tec, Data provided by Google Trends}
-	\label{fig:digdocker}	
-\end{figure}
-
-Figure \ref{fig:digdocker} shows the steadily rising interest in the most popular cloud technologies. 
-Docker is a container-engine enabling OS-Virtualization, Kubernetes is a Container-Platform manager and OpenShift is a software-suite based on Kubernetes. 
-These technologies will be summarized in \ref{subsec:Req}.
 
@@
 \end{figure}
-\newpage
 
@@
 This is one of the primary goals of the DevOps approach \cite{JA16}.
-\newpage
+
 \subsection{Sharing Results}
@@
 As a further example, the workspace could be shut down out of normal business-time. This attribute is crucial for a rented infrastructure, which is paid usually on both workload and storage.
-\subsection{Knowledge Required}
-The primary problem with Che is the requirement for an already dense knowledge of cloud related technology.
-Setting up the Workspace, docker-images, build and CI scripts is a complex topic and can not be done by someone getting started. 
-This requirement extends when working with a multi-pod-environment. 
-While most of the tasks are by themselves rather simple, the bare amount of such tasks (especially when migrating an existing project to Che) requires a broad knowledge.
-This can be done by a single expert user, enabling the other developers  to  use  the  build-routines and  environments provided.
 
@@
 
-\subsection{Vulnerability}
-The last problem is the vulnerability to malevolent users. 
-Someone who already is in possession of a valid account can drain the shared resources of a full cluster, therefore blocking  any  other  user  from  work.
-This  can  also  be achieved by human error. 
-It is therefore important to not only restrict access to project resources, but also the double-check and guard the releases.
 \section{Benefits for Education}
